---
title: "Regression models: mobility, intervention, and potential confounders"
author: "Kenneth Lee"
date: "15/09/2020"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: True
      smooth_scroll: True
---

## Data Preprocessing

```{r import packages, warning = FALSE, message = FALSE}
library(ggplot2)
library (readr)
library(tidyverse)
library(dplyr)
library(covidcast)
library(lubridate)
library(ggpubr)
library(reshape2)
library(tidyr)
library(viridis)
library(gridExtra)
library(zoo)
library(cowplot)
library(gplots)
library(car)

source("code/painter.r")
source("code/load_all_data.r")
source("code/shifter.r")
```

```{r define global variables}
STARTDATE <- "2020-02-20"
ENDDATE <- lubridate::today()
GEO_TYPE = "state" # state-level
GEO_VALUE = "*" # all states
EXCLUDED_AREAS = c("as","gu", "mp","vi") # excluded areas due to small sample size
```

```{r import data, warning = FALSE, message=FALSE}
data <- load_covidcast_data(STARTDATE, ENDDATE, GEO_TYPE, GEO_VALUE, EXCLUDED_AREAS)

# Read government intervention data
urlfile="https://raw.githubusercontent.com/COVID19StatePolicy/SocialDistancing/master/data/USstatesCov19distancingpolicy.csv"
policy <- read_csv(url(urlfile))

```

```{r preprocess policy data}
# Convert to lower case
policy$StatePostal <- tolower(policy$StatePostal)
# First we convert the date to a proper format
policy[, c("DateIssued", "DateEnacted", "DateExpiry" ,"DateEased", "DateEnded", "DateReexpanded1", "DateReeased1")] <- data.frame(lapply(policy[, c("DateIssued", "DateEnacted", "DateExpiry" ,"DateEased", "DateEnded", "DateReexpanded1", "DateReeased1")], function(x) as.Date(as.character(x), "%Y%m%d")))

# Get the dates between start and end date
all.dates <- seq(as.Date(STARTDATE), as.Date(ENDDATE), by="days")
time_value <- sort(rep(all.dates, length(unique(policy$StatePostal)) )) 

# Generate geo_value
geo_value <- rep(unique(policy$StatePostal), length(all.dates))
policy_signal <- data.frame(time_value = time_value, geo_value = geo_value)

# Create empty columns
policy_signal[,unique(policy$StatePolicy)] <- 0

# Fill in the count for each date
  # Get the policy name and state to filer policy signal 
for (row in (1:nrow(policy))){
  current.policy <- policy[row,]$StatePolicy
  current.state <- policy[row,]$StatePostal
  if (is.na(policy[row,]$DateEnded)){
    # Filter the rows of dataframe to be the current state and the time value that is after the policy is enacted.
    policy_signal[policy_signal$geo_value == current.state & policy_signal$time_value > as.Date(policy[row,]$DateEnacted), current.policy] <- 1
  }else{
    # Get time range between Date Enacted and Date Ended
    time.range <- seq(as.Date(policy[row,]$DateEnacted), as.Date(policy[row,]$DateEnded), by = "days")
    
    # Fill in the the rows that are in the current policy and fall between the time arrange to be 1
    policy_signal[policy_signal$time_value %in% time.range & policy_signal$geo_value == current.state, current.policy] <- 1
    }
}

# Compute the sum of the number of policies for every day in the state
policy_signal$total.num.policy <- rowSums(policy_signal[unique(policy$StatePolicy)])

# Compute the average on a 7day sliding window
policy_signal <-policy_signal %>%
    arrange(desc(geo_value)) %>% 
    group_by(geo_value) %>% 
    mutate(num.policy.7avg = rollmean(total.num.policy, k = 7, fill = NA))%>%
    ungroup()

# Finalize the covidcast-like signal for governemnt intervention
covidcast.like.policy.signal <- policy_signal %>% transmute(
  geo_value = geo_value,
  signal = "policy_7dav_num",
  time_value = time_value,
  direction = NA,
  issue = lubridate::today(),
  lag = issue - time_value,
  value = num.policy.7avg,
  stderr = NA,
  sample_size = NA,
  data_source = 'University of Washington')
```


```{r create a unified dataframe with mobility-intervention-confounders}
# We fist get all the signals
ftime.mobility <- data$Full.Time.Mobility[c("time_value","geo_value","value")] #mobility

case.count <- data$Avg.Confirmed.Case.Count[c("time_value","geo_value","value")]# case count

cumulative.case.count <-data$Cum.Avg.Case.Count[c("time_value","geo_value","value")]# cumulative case count

cumulative.case.count.prop <-data$Cum.Avg.Case.Count.Prop[c("time_value","geo_value","value")] # cumulative case count per 100,000

avg.death.count <- data$Avg.Death.Case.Count[c("time_value","geo_value","value")] #death count

cum.avg.death.count <- data$Cum.Avg.Death.Count[c("time_value","geo_value","value")] #cumulative death count

doc.visit <- data$smoothed_cli[c("time_value","geo_value","value")] # doctor visit

adj.doc.visit <- data$smoothed_adj_cli[c("time_value","geo_value","value")] # adjusted doctor visit

# Change the column name
colnames(ftime.mobility)[3]<-"full_time_work_prop"
colnames(case.count)[3] <- "confirmed_7dav_incidence_num"
colnames(cumulative.case.count)[3] <- "confirmed_7dav_cumulative"
colnames(cumulative.case.count.prop)[3] <- "confirmed_7dav_cumulative_prop"
colnames(avg.death.count)[3]<-"deaths_7dav_incidence_num"
colnames(cum.avg.death.count)[3] <- "deaths_7dav_cumulative_num"
colnames(doc.visit)[3] <- "smoothed_cli"
colnames(adj.doc.visit)[3] <- "smoothed_adj_cli"

# Create a list of confounders for left join with mobility
confounders <- list(case.count, cumulative.case.count, cumulative.case.count.prop,
     avg.death.count, cum.avg.death.count, doc.visit, adj.doc.visit)

# Turn the polical signal to be factors
factored.policy.signal <- cbind(policy_signal[1:2], lapply(policy_signal[3:17], as.factor),policy_signal[18:19])

# Intervention left join mobility with policy signal
intervention_mobility_case <- left_join(ftime.mobility,factored.policy.signal , by=c("time_value", "geo_value"))

# Left join again with all other potential confounders
for (confounder in confounders){
  intervention_mobility_case <- left_join(intervention_mobility_case, confounder, by=c("time_value", "geo_value"))
}
# Filter state "pr" as it is not available in the intervention data
factored_data <- intervention_mobility_case %>% filter(!(geo_value %in% c("pr")))

```

## Treatment effect of government interventions

Moreover, we may be interested to know the following:

* Is there a difference in the means of mobility signal across state?

* Is there a difference in the means of mobility signal in terms of stay-at-home order?

* Is there an interaction between the factor of states and the factor of the order?

To answer these questions, assuming that the data are normally distributed and the variance across groups are homogeneous, we will use ANOVA (Analysis of Variance). We will check these assumptions in the model diagnostics. 


### Main effect of States

We can see that some states have a particularly higher range of the mean mobility signal. For example, Montana (MT) clearly stands out from the rest, whereas Hawaii (HI) has a much lower mean mobility signal from Feb. to Sep. in 2020.

```{r check main effect, fig.width = 15, fig.height=10}
# Plotting the main effect of geo_value
plotmeans(full_time_work_prop~ geo_value,data=intervention_mobility_case, xlab="Geo_value", ylab="Mobility", main="Main effect (States)") 
```

### Interaction plots

```{r interaction plots across states}
for(i in 4:17){
  interaction.plot(factored_data[,"geo_value"], factored_data[,names(factored_data)[i]], factored_data$full_time_work_prop, xlab="States", ylab= "Full time work prop", main=paste("Interaction plot between states and", names(factored_data)[i], "on mobility"))  
}
```

### Boxplots of mobility across different levels of intervention

```{r boxplots for categorical variables, fig.height= 10, fig.width = 10}
# plot the boplots for categoical variables columns
p <- list()
counter <- 1
# loop through all categorical variables for intervention
for (i in 4:17){
p[[counter]]<- ggplot(factored_data, aes_string(x=names(factored_data)[3], y=names(factored_data)[i])) + 
    geom_boxplot() + 
     stat_summary(fun = mean, geom = "errorbar", aes(xmax = ..x.., xmin = ..x..), color ="red", linetype = 2)+
  labs(title=paste("Mobility by", names(factored_data)[i]), x="Mobility", y=names(factored_data)[i]) + theme(plot.title = element_text(size=9))

counter <- counter + 1
}
# Plot all the ggplot
do.call(grid.arrange,p)
```

### Distribution of mobility by various intervention across states

```{r mobility signal by intervention}
# grouped boxplot
for (i in 4:17){
  
p <- ggplot(factored_data, aes_string(x=names(factored_data)[3], y=names(factored_data)[2], fill=names(factored_data)[i])) + 
    geom_boxplot() +
  labs(title=paste("Mobility by State and", names(factored_data)[i]),x="mobility", y="State")

print(p)
}

```

### Transformations of the Response Variable

```{r visualize all histogram of reponse}
par(mfrow=c(2,2))
hist(factored_data$full_time_work_prop, main="Histogram of unchanged response", xlab="full_time_work_prop")
hist(log(factored_data$full_time_work_prop), main="Histogram of log response", xlab="full_time_work_prop")
hist(1/(factored_data$full_time_work_prop), main="Histogram of 1/response", xlab="full_time_work_prop")
hist(sqrt(factored_data$full_time_work_prop), main="Histogram of squared root of response", xlab="full_time_work_prop")
```

### Histograms of all continuous variables

```{r visualize the distributions of the continuous variables, fig.height= 10, fig.width = 10}
par(mfrow=c(3,3))
for (i in 19:ncol(factored_data)) 
{
  hist(factored_data[,names(factored_data)[i]], xlab=names(factored_data)[i], main=paste("Histogram of", names(factored_data)[i]))
}
par(mfrow=c(1,1))
```

### Linear regression models comparsion (Mobility, Intervention, Confounders)

We modified the data so that the confounders are shifted n days forwarded in time based on our initial exploration. 

```{r linear regression model}
# select best lag number based on intial exploration
case.shifted.days.spearman <- 26 # based spearman correlation
doc.visit.shifted.days.spearman <- 33 # based spearman correlation for doctor visit
cum.death.shifted.days.spearman <- 26
death.shifted.days.spearman <- 100

case.vec <- c("confirmed_7dav_incidence_num", "confirmed_7dav_cumulative", "confirmed_7dav_cumulative_prop")
doc.vec <- c("smoothed_cli", "smoothed_adj_cli")
cum.death.vec <- c("deaths_7dav_cumulative_num")
death.vec <- c("deaths_7dav_incidence_num")

for (state in unique(factored_data$geo_value)){
  
  # Filter dataframe by every state
  selected.df <- factored_data %>% filter(geo_value==state)
  
  # Shift the case count column vector by the specified shift time
  factored_data <- shiftDays(selected.df, factored_data, case.shifted.days.spearman, case.vec)
  # Shift the  doctor column vector by the specified shift time
  factored_data <-shiftDays(selected.df, factored_data, doc.visit.shifted.days.spearman, doc.vec)
  
  # Shift the death count column vector by the specified shift time
  factored_data <-shiftDays(selected.df, factored_data, cum.death.shifted.days.spearman, cum.death.vec)
  # Shift the cum death count column vector by the specified shift time
  factored_data <-shiftDays(selected.df, factored_data, death.shifted.days.spearman, death.vec)
}


#Problematic: EmergDec, deaths_7dav_incidence_num

# model 1 : y is mobility, x's are the confounders and intervention
# Note: we dropped the geo and time value

lm.fit.1 <- lm(full_time_work_prop ~ confirmed_7dav_incidence_num+
                 confirmed_7dav_cumulative+
                 confirmed_7dav_cumulative_prop+
                 PublicMask+
                 StayAtHome+
                 SchoolClose+
                 BarRestrict+
                 RestaurantRestrict+
                 Quarantine+
                 NEBusinessClose+
                 TravelRestrictIntra+
                 TravelRestrictExit+
                 TravelRestrictEntry+
                 GathRecom+
                 total.num.policy+
                 num.policy.7avg+
                deaths_7dav_cumulative_num+
                 smoothed_cli+
                 smoothed_adj_cli+
                 CaseIsolation, data=factored_data[,-c(1,2,4,24)], na.action = na.exclude)

summary(lm.fit.1)
# Plot the normal qq plot, residuals vs. fitted values
par(mfrow=c(2,2))
plot(lm.fit.1)

# model 1: include all two way interaction 
lm.fit.2 <- lm(full_time_work_prop~.^2, data=factored_data[,-c(1,2,4,24)])
summary(lm.fit.2)
plot(lm.fit.2)
par(mfrow=c(1,1))
```


### ANOVA

### Interpret the result

```{r ANOVA table}
anova(lm.fit.1)

anova(lm.fit.2)


# Change the default setting for unbalanced ANOVA
#options(contrasts = c("contr.sum", "contr.poly"))

# ANOVA test for unbalanced designs.

# print the ANOVA table for both models
#Anova(lm.fit.0, type = "III")

#Anova(lm.fit.1, singular.ok = T, type = "III") #set singular.ok = T to avoid error 
```

### Model selection

```{r model validation}

```
  
  




