---
title: "Estimating Causal Effect of Government Intervention to Mobility"
author: "Kenneth Lee"
date: "23/09/2020"
output: html_document
---

```{r import packages, echo=FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library (readr)
library(tidyverse)
library(dplyr)
library(covidcast)
library(lubridate)
library(ggpubr)
library(reshape2)
library(tidyr)
library(viridis)
library(gridExtra)
library(data.table)
library(zoo)
library(cowplot)
library(CausalImpact)

source("code/painter.r")
source("code/load_all_data.r")
```

```{r define global variables, echo=FALSE}
STARTDATE <- "2019-01-01"
ENDDATE <- lubridate::today()
GEO_TYPE = "state" # state-level
GEO_VALUE = "ca" # all states
#EXCLUDED_AREAS = c("as","gu", "mp","vi") # excluded areas due to small sample size
DT_X = 7 # 	 Time shifts to consider for x

```

```{r import data, warning = FALSE, message=FALSE, echo=FALSE}
# Get mobility signal
ftime_mobility <- covidcast_signal(data_source = "safegraph", signal ="full_time_work_prop",start_day = STARTDATE, end_day = ENDDATE,geo_type = GEO_TYPE, geo_values = GEO_VALUE)

# Create a dataframe with mobility and other potential covariates
tbl <-list.files(path = "data/temperature",
               pattern = "*.csv", 
               full.names = T) %>% map_df(~read_csv(.)) 

intervention <- load_policy_data(STARTDATE, ENDDATE)

# US holidays
ca.holidays <- c("2019-01-01", "2019-01-21", "2019-02-18", "2019-05-27", "2019-07-04", "2019-09-02", "2019-10-14", "2019-11-11", "2019-11-28", "2019-12-25", "2019-04-01", "2020-01-01", "2020-01-20", "2020-02-17", "2020-05-25", "2020-07-03", "2020-09-07", "2020-10-12", "2020-11-11", "2020-11-26", "2020-12-25")

# Convert data type
ca.holidays <-sapply(ca.holidays, as.Date)
```

## Introduction

We would like to estimate the causal effect of governement intervention, collected from University of Washington, on the mobility signal, produced by covidcast signal by using Baysiean structural time-series model. The question we want to answer is: how the mobility signal was increased/decreased by various state policies? 

Particularly, we will use the R package called [`CausalImpact`](https://google.github.io/CausalImpact/CausalImpact.html). This package constructs a Bayesian structural time-series model by using a given response time series (e.g. mobility signal) and a set of control time-series (e.g. mobility signal in state that doesn't have the policy). We use this package because 1.) our data is not a randomized experiement 2.) 

## Model assumption

* We assume that there is a set control time series that were themselves not affected by the intervention. 

* The model also assumes that the relationship between covariates and treated time series, as established during the pre-period, remains stable throughout the post-period.

## Implementation

* We will use the full-time mobility signal as the treated time series. We then select the following as our control time-series: average temperature, holiday (binary), weekend indicator (binary).

```{r preprocess, echo=FALSE}

# Get the average temperature
avg.temp <- tbl %>%
  group_by(DATE) %>%
  summarize(avg.temp = mean(TAVG, na.rm=TRUE))
colnames(avg.temp)[1] <- "time_value"

# Get mobility data
mobility <- ftime_mobility[c("time_value", "value")]

# full outer join 
df <- full_join(mobility,avg.temp, by="time_value")

# Get the last time stamp available in data
lastdate <- tail(df$time_value, n=1)
# Create dummy indicators 
weekend <- ifelse(weekdays(seq(as.Date(STARTDATE), as.Date(lastdate), by="days")) %in% c("Saturday", "Sunday"), 1, 0)
df$weekend <- weekend

# create dummy variables for holidays
holidays <- ifelse(seq(as.Date(STARTDATE), as.Date(lastdate), by="days") %in% ca.holidays, 1, 0)
df$holiday <- holidays

# get intervention data for only ca
ca_policy <- intervention %>% filter(geo_value == "ca" & total.num.policy > 0)

# get the first date of the intervention
interventio.first.day <- ca_policy$time_value[1]

# Create pre and post period for the causal impact package
pre.period <- as.Date(c(STARTDATE, as.character(interventio.first.day-1)))
post.period <- as.Date(c(interventio.first.day, ENDDATE))


data <- zoo(cbind(df$value, df$avg.temp, df$weekend), df$time_value)

# plot the see how mobility may correlate with average temperature of CA
matplot(cbind(log(df$value), log(df$avg.temp), df$weekend), type = "l")
```

### Visualize the causal estimate 

We show that the causal estimate by counterfactual prediction (blue line) and observations (black line)

```{r build model}
impact <- CausalImpact(data, pre.period, post.period)
plot(impact) + labs(title="Causal estimate by bayesian structual time-series")
```

### Interpret the result 
```{r generate_report, echo=FALSE}
summary(impact)
summary(impact, "report")
```
### Reference

* “CausalImpact 1.2.1, Brodersen et al., Annals of Applied Statistics (2015). http://google.github.io/CausalImpact/”

