---
title: "Estimating Causal Effect of Government Intervention to Mobility via Regression Discontinuity"
author: "Kenneth Lee"
date: "23/09/2020"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: True
      smooth_scroll: True
---

```{r import packages, echo=FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library (readr)
library(tidyverse)
library(dplyr)
library(covidcast)
library(lubridate)
library(ggpubr)
library(reshape2)
library(tidyr)
library(viridis)
library(gridExtra)
library(data.table)
library(zoo)
library(cowplot)
library(CausalImpact)

source("code/painter.r")
source("code/load_all_data.r")
source("code/shifter.r")
```

```{r define global variables, echo=FALSE}
STARTDATE <- "2019-01-01"
ENDDATE <- lubridate::today()
GEO_TYPE = "state" # state-level
GEO_VALUE = "ca" # all states
EXCLUDED_AREAS = c("as","gu", "mp","vi") # excluded areas due to small sample size
DT_X = 7 # 	 Time shifts to consider for x

```

```{r import data, warning = FALSE, message=FALSE, echo=FALSE}
# Get the covidcast data accordingly
data <- load_covidcast_data(STARTDATE, ENDDATE, GEO_TYPE, GEO_VALUE, EXCLUDED_AREAS)

intervention <- load_policy_data(STARTDATE, ENDDATE)

# Filter to only specified one state
all.ca.policy <- intervention %>% filter(geo_value == GEO_VALUE)

# We fist get all the signals
ftime.mobility <- data$Full.Time.Mobility[c("time_value","value")] #mobility

case.count <- data$Avg.Confirmed.Case.Count[c("time_value","value")]# case count

cumulative.case.count <-data$Cum.Avg.Case.Count[c("time_value","value")]# cumulative case count

cumulative.case.count.prop <-data$Cum.Avg.Case.Count.Prop[c("time_value","value")] # cumulative case count per 100,000

avg.death.count <- data$Avg.Death.Case.Count[c("time_value","value")] #death count

cum.avg.death.count <- data$Cum.Avg.Death.Count[c("time_value","value")] #cumulative death count

doc.visit <- data$smoothed_cli[c("time_value","value")] # doctor visit

adj.doc.visit <- data$smoothed_adj_cli[c("time_value","value")] # adjusted doctor visit
```


```{r preprocess, echo=FALSE, warning = FALSE, message=FALSE, fig.height= 15, fig.width = 15}
# Change the column name
colnames(ftime.mobility)[2]<-"full_time_work_prop"
colnames(case.count)[2] <- "confirmed_7dav_incidence_num"
colnames(cumulative.case.count)[2] <- "confirmed_7dav_cumulative"
colnames(cumulative.case.count.prop)[2] <- "confirmed_7dav_cumulative_prop"
colnames(avg.death.count)[2]<-"deaths_7dav_incidence_num"
colnames(cum.avg.death.count)[2] <- "deaths_7dav_cumulative_num"
colnames(doc.visit)[2] <- "smoothed_cli"
colnames(adj.doc.visit)[2] <- "smoothed_adj_cli"

# Create a list of confounders for left join with mobility
confounders <- list(case.count, cumulative.case.count, cumulative.case.count.prop,
     avg.death.count, cum.avg.death.count, doc.visit, adj.doc.visit)

# Intervention left join mobility with policy signal
intervention_mobility_case <- left_join(ftime.mobility, all.ca.policy, by=c("time_value"))

# Left join again with all other potential confounders
for (confounder in confounders){
  intervention_mobility_case <- left_join(intervention_mobility_case, confounder, by=c("time_value"))
}


# get intervention data for only ca
ca_policy.valid.time <- intervention %>% filter(geo_value == "ca" & total.num.policy > 0)

# get the first date of the intervention
intervention.first.day <- ca_policy.valid.time$time_value[1]

# Plot the time-series: mobility signal as y, time as x 
intervention_mobility_case %>%
ggplot(aes(x = time_value, y = full_time_work_prop)) + 
  geom_point() +
  geom_vline(xintercept = intervention.first.day) + 
  labs(y = "Full time away home signal", x = "time-value")
```

## Lag analysis at a state-level

```{r state-lag-analysis, echo=FALSE, warning = FALSE, message=FALSE, fig.height= 15, fig.width = 15}

SHIFTDAY <- 100
corr.method <- "spearman"
by_method <- "geo_value"
title <- "Median Rank correlation between other signals and future mobility (slicing by state)"

covidcastlike.signals <- list(
  data$Avg.Confirmed.Case.Count, data$Cum.Avg.Case.Count, data$Cum.Avg.Case.Count.Prop, data$Avg.Death.Case.Count, data$Cum.Avg.Death.Count, data$smoothed_cli, data$smoothed_adj_cli)

names <- list("7-day avg. confirmed case", 
             "Cum 7day avg. confirmed case",
             "Cum 7day avg. confirmed case, per 100,000",
             "death case",
             "cumulative death case",
             "doctor visit",
             "doctor visit (day-of-week effects removed)")


# Compute pearson correlation between other covidcast-like signals and mobility
plot.all.Corr.Median.by.shift(covidcastlike.signals, 
                              data$Full.Time.Mobility,
                              SHIFTDAY,
                              names, 
                              corr.method, 
                              title, 
                              by_method)
```

## Use most correlated shift time to build linear models

$$y = \beta_{0} + \beta_{1}(\sum I(A_i = 1)) + \beta_{2}x_1 + \beta_{3}x_2$$


```{r linear regression}
# select best lag number based on intial exploration
case.shifted.days.spearman <- 37 # based spearman correlation
doc.visit.shifted.days.spearman <- 62 # based spearman correlation for doctor visit
cum.death.shifted.days.spearman <- 55
death.shifted.days.spearman <- 55

case.vec <- c("confirmed_7dav_incidence_num", "confirmed_7dav_cumulative", "confirmed_7dav_cumulative_prop")
doc.vec <- c("smoothed_cli", "smoothed_adj_cli")
cum.death.vec <- c("deaths_7dav_cumulative_num")
death.vec <- c("deaths_7dav_incidence_num")

# Make two copies for shifting the data
selected.df <- intervention_mobility_case
factored_data <- intervention_mobility_case
# Change the data by shifting the covariates
# Shift the case count column vector by the specified shift time
factored_data <- shiftDays(selected.df, factored_data, case.shifted.days.spearman, case.vec)

# Shift the  doctor column vector by the specified shift time
factored_data <-shiftDays(selected.df, factored_data, doc.visit.shifted.days.spearman, doc.vec)
  
# Shift the death count column vector by the specified shift time
factored_data <-shiftDays(selected.df, factored_data, cum.death.shifted.days.spearman, cum.death.vec)

# Shift the cum death count column vector by the specified shift time
factored_data <-shiftDays(selected.df, factored_data, death.shifted.days.spearman, death.vec)
  
# We specifically look at emergency declaration
factored_data%>% 
  mutate(EmergDec.duration = cumsum(EmergDec)) %$% 
  lm(full_time_work_prop ~ EmergDec.duration + smoothed_cli+smoothed_adj_cli ) %>% 
  summary()
```

```{r more linear models}
# We specifically look at emergency declaration
factored_data%>% 
  mutate(EmergDec.duration = cumsum(EmergDec)) %$% 
  lm(full_time_work_prop ~ EmergDec.duration + smoothed_cli+smoothed_adj_cli+confirmed_7dav_cumulative_prop+deaths_7dav_cumulative_num) %>% 
  summary()
```

## Model with shifting the covariates (weekends included)

```{r plot the regression, echo=FALSE, warning = FALSE, message=FALSE, fig.height= 15, fig.width = 15}
# Create a new dataframe
factored_data.lm <- factored_data%>% 
  mutate(EmergDec.duration = cumsum(EmergDec)) 

# Fit a linear regression model
lm.fit <- lm(full_time_work_prop ~ EmergDec.duration + smoothed_cli+smoothed_adj_cli, data =factored_data.lm)  

# Pad the fitted values with NA 
factored_data.lm$predlm <- c(rep(NA, 458),predict(lm.fit))

# Plot the graph
p <- factored_data.lm %>% 
  mutate(policy.duration = cumsum(EmergDec), EmergDeclaration = as.factor(EmergDec)) %>% 
  ggplot(aes(x = time_value, y = full_time_work_prop, color = EmergDeclaration )) +
  geom_point() + 
  geom_line(aes(x = time_value, y = predlm, colour="fitted value"), size = 1) +
   labs(title = "Covariates selected WITH most correlated number of shift")
  #geom_smooth(method = "lm")
p

```

## Model without shifting the covariates (weekends included)

```{r LR model without selecting most correlated lag covariates, warning = FALSE, message=FALSE, fig.height= 15, fig.width = 15}
intervention.lm <- intervention_mobility_case %>% 
  mutate(EmergDec.duration = cumsum(EmergDec)) 

lm.fit.no.lag <- lm(full_time_work_prop ~ EmergDec.duration + smoothed_cli+smoothed_adj_cli, data =intervention.lm
) 

intervention.lm$predlm <- c(rep(NA, nrow(intervention.lm) - length(predict(lm.fit.no.lag))), predict(lm.fit.no.lag))

intervention.lm%>% 
  mutate(policy.duration = cumsum(EmergDec), EmergDeclaration = as.factor(EmergDec)) %>% 
  ggplot(aes(x = time_value, y = full_time_work_prop, color = EmergDeclaration)) +
  geom_point() + 
  geom_line(aes(x = time_value, y = predlm, colour="fitted value"), size = 1)+
  labs(title = "Covariates selected WITHOUT most correlated number of shift")
```

## Weekend effects

We suspect that the mobility signal is lower than usual during the weekend.

```{r examine weekend effects from the data}
intervention_mobility_case$weekday <- weekdays(as.Date(intervention_mobility_case$time_value)) 

p <- ggplot(intervention_mobility_case, aes(x=weekday, y=full_time_work_prop)) + 
  geom_boxplot()
p
```


## Re-plot the regression line after dropping weekends

### Covariates shifted

```{r plot the regression line for the data WITH shift after dropping weekends, echo=FALSE, warning = FALSE, message=FALSE, fig.height= 15, fig.width = 15}
# Get lm summary to compare
factored_data.without.weekend <- factored_data%>% 
  mutate(EmergDec.duration = cumsum(EmergDec), EmergDeclaration = as.factor(EmergDec), weekday= weekdays(as.Date(time_value)))%>% 
  filter(!weekday %in% c("Saturday", "Sunday")) 

# Check the lm.fit
factored_data.without.weekend %$%
  lm(full_time_work_prop ~ EmergDec.duration + smoothed_cli+smoothed_adj_cli ) %>% 
  summary()

# Get all the fitted values
shifted.predlm <- factored_data.without.weekend%$%
  lm(full_time_work_prop ~ EmergDec.duration + smoothed_cli+smoothed_adj_cli) %>%
  predict()

# Pad the fitted values with NA 
factored_data.without.weekend$predlm <- c(rep(NA, nrow(factored_data.without.weekend) - length(shifted.predlm)), shifted.predlm)

factored_data.without.weekend %>%
  ggplot(aes(x = time_value, y = full_time_work_prop, color = EmergDeclaration)) +
  geom_point() + 
  geom_line(aes(y = predlm, colour="fitted value"), size = 1) +
   labs(title = "Covariates selected WITH most correlated number of shift (weekends dropped)")
```


### Covariates not shifted


```{r plot the regression line for the data WITHOUT shift after dropping weekends, echo=FALSE, warning = FALSE, message=FALSE, fig.height= 15, fig.width = 15}

# Filter the dataset
noshift.without.weekends <-intervention_mobility_case%>%
  mutate(policy.duration = cumsum(EmergDec), EmergDeclaration = as.factor(EmergDec), weekday = weekdays(as.Date(time_value))) %>% 
  filter(!weekday %in% c("Saturday", "Sunday"))
  
# Print lm summary
 noshift.without.weekends %$%
  lm(full_time_work_prop ~ policy.duration + smoothed_cli+smoothed_adj_cli) %>% summary
  
# Get the fitted value  
noshift.predlm <- noshift.without.weekends %$%
  lm(full_time_work_prop ~ policy.duration + smoothed_cli+smoothed_adj_cli) %>%
  predict()

# Pad the fitted values with NA 
noshift.without.weekends$predlm <- c(rep(NA, nrow(noshift.without.weekends) - length(noshift.predlm)), noshift.predlm)

# Plot the graph
noshift.without.weekends %>%
  ggplot(aes(x = time_value, y = full_time_work_prop, color = EmergDeclaration)) +
  geom_point() + 
  geom_line(aes(y = predlm, colour="fitted value"), size = 1)+
  labs(title = "Covariates selected WITHOUT most correlated number of shift (weekends dropped)")
```