---
title: "Ranking the effect of intervention at countylevel"
author: "Kenneth Lee"
date: "06/01/2021"
output:
  html_document:
    code_folding: hide
    keep_md: yes
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Abstract

This notebook has two objectives. Firstly, it aims to measure the effect of interventions on mobility with other potential confounding signals such as confirmed case counts. The motivation is to observe whether the effect of intervention is greater than the effect of case signals on mobility. Secondly, we will rank the effects of various government interventions.

Due to the limitation of computation, we focus on analyzing the state of California and its counties. We will regress mobility signals on the covariates of interests and measure the effect of the covariates based on its regression coefficients.

```{r load_data, message=F, warning=F}
library(dplyr)
library(lubridate)
library(covidcast)
library(RcppRoll)
library(tidycensus)
library(reshape2)
library(zoo)
library(readxl)
library(ggplot2)
library(gridExtra)


source("code/loader.r")
source("code/parser.r")

# Load files 
confirmed_7dav_cumulative_prop <- read_excel("data/case_signals/ca-county-confirmed_7dav_cumulative_prop.xlsx")
 
confirmed_7dav_incidence_prop <- read_excel("data/case_signals/ca-county-confirmed_7dav_incidence_prop.xlsx")
 
deaths_7dav_cumulative_prop <- read_excel("data/case_signals/ca-county-deaths_7dav_cumulative_prop.xlsx")

deaths_7dav_incidence_prop <- read_excel("data/case_signals/ca-county-deaths_7dav_incidence_prop.xlsx")

tx.confirmed_7dav_cumulative_prop <- read_excel("data/case_signals/tx.confirmed_7dav_cumulative_prop.xlsx")

tx.confirmed_7dav_incidence_prop <- read_excel("data/case_signals/tx.confirmed_7dav_incidence_prop.xlsx")

tx.deaths_7dav_cumulative_prop <- read_excel("data/case_signals/tx.deaths_7dav_cumulative_prop.xlsx")


tx.deaths_7dav_incidence_prop <- read_excel("data/case_signals/tx.deaths_7dav_incidence_prop.xlsx")

STARTDATE <- "2020-02-20"
ENDDATE <- "2020-12-15"


fipscodes <- fips_codes%>%
  mutate(fips =paste(state_code, county_code, sep=""))

# Get counties names
ca.counties <- fipscodes%>%
  filter(state=="CA")%>% select(county)
ca.counties <- ca.counties$county

tx.counties <- fipscodes%>%
  filter(state=="TX")%>% select(county)
tx.counties <- tx.counties$county

ca.rest <- read.csv("data/ca.rest.csv",stringsAsFactors = F)

ca.chome <- read.csv("data/ca.chome.csv",stringsAsFactors = F)


tx.rest <- read.csv("data/tx.rest.csv",stringsAsFactors = F)

tx.chome <- read.csv("data/tx.chome.csv",stringsAsFactors = F)

tx.chome$time_value <- as.Date(tx.chome$time_value)
tx.rest$time_value <- as.Date(tx.rest$time_value)
ca.chome$time_value <- as.Date(ca.chome$time_value)
ca.rest$time_value <- as.Date(ca.rest$time_value)

ca.rest$geo_value <- paste("0",ca.rest$geo_value, sep="")
ca.chome$geo_value <- paste("0",ca.chome$geo_value, sep="")
tx.chome$geo_value <- as.character(tx.chome$geo_value)
tx.rest$geo_value <- as.character(tx.rest$geo_value)

# Convert the time_value to DATE
 confirmed_7dav_cumulative_prop$time_value <- as.Date(confirmed_7dav_cumulative_prop$time_value)
 
confirmed_7dav_incidence_prop$time_value <- as.Date(confirmed_7dav_incidence_prop$time_value)
 
deaths_7dav_cumulative_prop$time_value <- as.Date(deaths_7dav_cumulative_prop$time_value)
 
deaths_7dav_incidence_prop$time_value <- as.Date(deaths_7dav_incidence_prop$time_value)
# time value
tx.confirmed_7dav_cumulative_prop$time_value <- as.Date(tx.confirmed_7dav_cumulative_prop$time_value)

tx.confirmed_7dav_incidence_prop$time_value <- as.Date(tx.confirmed_7dav_incidence_prop$time_value)
 
tx.deaths_7dav_cumulative_prop$time_value <- as.Date(tx.deaths_7dav_cumulative_prop$time_value)
 
tx.deaths_7dav_incidence_prop$time_value <- as.Date(tx.deaths_7dav_incidence_prop$time_value)


# Convert the county fips codes to character

tx.confirmed_7dav_cumulative_prop$geo_value <- as.character(tx.confirmed_7dav_cumulative_prop$geo_value)
 
tx.confirmed_7dav_incidence_prop$geo_value <- as.character(tx.confirmed_7dav_incidence_prop$geo_value)
 
tx.deaths_7dav_cumulative_prop$geo_value <- as.character(tx.deaths_7dav_cumulative_prop$geo_value)
 
tx.deaths_7dav_incidence_prop$geo_value <- as.character(tx.deaths_7dav_incidence_prop$geo_value)
 



# Pad 0 to all CA data
confirmed_7dav_cumulative_prop$geo_value <- paste("0", confirmed_7dav_cumulative_prop$geo_value, sep="" )
 
confirmed_7dav_incidence_prop$geo_value <- paste("0", confirmed_7dav_incidence_prop$geo_value, sep="" )
 
deaths_7dav_cumulative_prop$geo_value <- paste("0", deaths_7dav_cumulative_prop$geo_value, sep="" )
 
deaths_7dav_incidence_prop$geo_value <- paste("0", deaths_7dav_incidence_prop$geo_value, sep="" )
 


# Policy data
policy <- load_policy()
ca.policy <- policy %>% filter(StatePostal == "ca", StateWide == 1)
tx.policy <- policy %>% filter(StatePostal == "tx", StateWide == 1)

ca.rest$county <- fips_to_name(ca.rest$geo_value)
ca.chome$county <- fips_to_name(ca.chome$geo_value)
tx.rest$county <- fips_to_name(tx.rest$geo_value)
tx.chome$county <- fips_to_name(tx.chome$geo_value)


ca.policy_signal<- getSumOfPolicy(ca.policy , STARTDATE, ENDDATE)
tx.policy_signal<- getSumOfPolicy(tx.policy , STARTDATE, ENDDATE)

# Turn the polical signal to be factors
factored.ca.policy.signal <- cbind(ca.policy_signal[1], 
                                lapply(ca.policy_signal[3:14],
                                       as.factor),
                                ca.policy_signal[15:16])


ca.holidays <- c("2019-01-01",
                 "2019-01-21",
                 "2019-02-18",
                 "2019-04-01",
                 "2019-05-27",
                 "2019-07-04",
                 "2019-09-02",
                 "2019-09-09",
                 "2019-09-27",
                 "2019-11-11",
                 "2019-11-28",
                 "2019-12-25",
                 "2020-01-01", 
                 "2020-01-20",
                 "2020-02-17",
                 "2020-03-31",
                 "2020-05-25",
                 "2020-07-03",
                 "2020-09-07",
                 "2020-09-09",
                 "2020-09-25",
                 "2020-11-11",
                 "2020-11-26",
                 "2020-12-25",
                 "2021-01-01",
                 "2021-01-18")
tx.holidays <- c("2019-01-01",
                 "2019-01-19",
                 "2019-01-21",
                 "2019-02-18",
                 "2019-03-02",
                 "2019-03-31",
                 "2019-04-19",
                 "2019-04-21",
                 "2019-05-27",
                 "2019-06-19",
                 "2019-07-04",
                 "2019-09-02",
                 "2019-11-11",
                 "2019-11-28",
                 "2019-12-24",
                 "2019-12-25",
                 "2019-12-26",
                 "2020-01-01",
                 "2020-01-19",
                 "2020-01-20",
                 "2020-02-17",
                 "2020-03-02",
                 "2020-03-21",
                 "2020-04-10",
                 "2020-04-21",
                 "2020-05-25",
                 "2020-06-19",
                 "2020-07-03",
                 "2020-09-07",
                 "2020-11-11",
                 "2020-11-26",
                 "2020-12-24",
                 "2020-12-25",
                 "2020-12-26")

# May through October is wildfire season
ca.wildfire_seasons <- seq(as.Date("2020-05-01"), as.Date("2020-10-31"), 1)

tx.wildfire_seasons <- seq(as.Date("2020-03-01"), as.Date("2020-05-31"), 1)

# airquality

ca.air.2019 <- read.csv("data/air.quality/ad_viz_plotval_data_pm2.5_2019.csv")

ca.air.2020 <- read.csv("data/air.quality/ad_viz_plotval_data_pm2.5_2020.csv")

ca.air.2021 <- read.csv("data/air.quality/ad_viz_plotval_data_pm2.5_2021.csv")



tx.air.2019 <- read.csv("data/air.quality/ad_viz_plotval_data_tx_2019.csv")

tx.air.2020 <- read.csv("data/air.quality/ad_viz_plotval_data_tx_2020.csv")

tx.air.2021 <- read.csv("data/air.quality/ad_viz_plotval_data_tx_2021.csv")

ca.air <- rbind(ca.air.2019, ca.air.2020, ca.air.2021)
tx.air <- rbind(tx.air.2019, tx.air.2020, tx.air.2021)
sel_col <- c("Date", "DAILY_AQI_VALUE","COUNTY")
ca.air <- ca.air[sel_col]
tx.air <- tx.air[sel_col]
# Change the date format
ca.air$Date <- as.Date(format(strptime(as.character(ca.air$Date), "%m/%d/%Y"), "%Y-%m-%d"))

tx.air$Date <- as.Date(format(strptime(as.character(tx.air$Date), "%m/%d/%Y"), "%Y-%m-%d"))

ca.air$COUNTY <- paste(ca.air$COUNTY, "County", sep=" ")
tx.air$COUNTY <- paste(tx.air$COUNTY, "County", sep=" ")
ca.air$DAILY_AQI_VALUE <- as.numeric(ca.air$DAILY_AQI_VALUE)
tx.air$DAILY_AQI_VALUE <- as.numeric(tx.air$DAILY_AQI_VALUE)


colnames(ca.air)[1] <- "time_value"
colnames(ca.air)[3] <- "county"

colnames(tx.air)[1] <- "time_value"
colnames(tx.air)[3] <- "county"

# Take the median AQI
ca.median_aqi <- ca.air %>%
  group_by(county, time_value)%>%
  summarize(median_AQI = median(DAILY_AQI_VALUE))

tx.median_aqi<- tx.air %>%
  group_by(county, time_value)%>%
  summarize(median_AQI = median(DAILY_AQI_VALUE))

```

```{r helper-function, message=F, warning=F}
################ getForwardDays () ###################
# compute the number of days within a given day range that
# gives the highest spearman correlation between df1 and df2

# argument
# df1: a dataframe 
# df2: a dataframe that the data needs to be forwarded
# dt_vec: a vector contains a range of integers

getForwardDays <- function(df1, df2, dt_vec){
  # Empty list
  df_list <- vector("list", length(dt_vec))
  
  for (i in 1:length(dt_vec)) {
    df_list[[i]] <- covidcast_cor(df1, df2, dt_y = dt_vec[i],
                                  by = "geo_value", 
                                  method="spearman")

    df_list[[i]]$dt <- dt_vec[i]
  }
  
  # Stack into one big data frame
  df <- do.call(rbind, df_list)
  

  return(df)
}

################ joinConfounder() ###################

joinConfounder <- function(mobility.df, policy.df, state="ca"){
  if(state=="ca"){
    # Combine case and mobility
  cum.confirmed.case <- confirmed_7dav_cumulative_prop[c("time_value","geo_value","value")]
  new.confirmed.case <- confirmed_7dav_incidence_prop[c("time_value","geo_value","value")]
  cum.death.case <- deaths_7dav_cumulative_prop[c("time_value","geo_value","value")]
  new.death.case <- deaths_7dav_incidence_prop[c("time_value","geo_value","value")]
  }else{
    # Combine case and mobility
    cum.confirmed.case <- tx.confirmed_7dav_cumulative_prop[c("time_value","geo_value","value")]
  new.confirmed.case <- tx.confirmed_7dav_incidence_prop[c("time_value","geo_value","value")]
  cum.death.case <- tx.deaths_7dav_cumulative_prop[c("time_value","geo_value","value")]
  new.death.case <- tx.deaths_7dav_incidence_prop[c("time_value","geo_value","value")]
  }

  colnames(cum.confirmed.case)[3] <- "confirmed_7dav_cumulative_prop"
  colnames(new.confirmed.case)[3] <- "confirmed_7dav_incidence_prop"
  colnames(cum.death.case)[3] <- "deaths_7dav_cumulative_prop"
  colnames(new.death.case)[3] <- "deaths_7dav_incidence_prop"
  
  # A list of confounders
  confounders <- list(cum.confirmed.case, 
                      new.confirmed.case, 
                      cum.death.case,
                      new.death.case)
  
  # Left join again with all other potential confounders
  for (confounder in confounders){
    mobility.df <- left_join(mobility.df, confounder, by=c("time_value", "geo_value"))
  }

  # combine with policy signal
  mobility.df <- left_join(mobility.df, policy.df, by=c("time_value"))
  
  if(state=="ca"){
    mobility.df <- left_join(mobility.df, ca.median_aqi, by=c("time_value", "county"))
  }else{
    mobility.df <- left_join(mobility.df, tx.median_aqi, by=c("time_value", "county"))
  }
  return(mobility.df)

}

################ prepDF() ###################

prepDF <- function(mobility.df, 
                     policy, 
                     states,
                     interventions,
                     fips_codes,
                     forward=F){
  
  # Convert to lower case
  lowerStates <- tolower(states)
  
  # filter policy
  policy.df <- policy %>% filter(StatePostal%in%lowerStates, StateWide == 1)
  
  policy.df <- getSumOfPolicy(policy.df, STARTDATE, ENDDATE)
  
  # convert to factor
  policy.df[interventions] <- lapply(policy.df[interventions], as.factor)

  # join the mobility and policy 
  df<- joinConfounder(mobility.df, policy.df, state = lowerStates)
  
  # Change the value name for fitting regression model
  colnames(df)[colnames(df)=="value"] <- unique(mobility.df$signal)
  return(df)
}

############ plot_bar() ###################

plot_bar <- function(df,
                     x,
                     y,
                     lower.bound,
                     upper.bound,
                     gp,
                     name){
  p<- ggplot(df, aes_string(x=x, y=y, group=gp, color=gp)) + 
  #geom_line() +
  geom_point()+
  geom_errorbar(aes_string(xmin=lower.bound, xmax=upper.bound), width=.2,
                 position=position_dodge(0.05))+
    ggtitle(name)
      
  return(p)
}

```

# Model:

$$Y_t = \beta_{0}+ \sum_{i}\beta_{i}S_{i_{t}} + \sum_{k}\beta_{k}P_{k_{t}}$$


## Completely stay at home: CA

```{r model, warning=F}
interventions <- c("EmergDec",
                   "GathRestrict",
                   "SchoolClose",
                   "BarRestrict",
                   "PublicMask",
                   "OtherBusinessClose",
                   "BusinessMask",
                   "SchoolMask",
                   "Quarantine")

confounders.names <- c("confirmed_7dav_cumulative_prop",
                  "confirmed_7dav_incidence_prop",
                  "deaths_7dav_cumulative_prop",
                  "deaths_7dav_incidence_prop")

mobility.name <- "completely_home_prop"

formula1 <- as.formula(paste(mobility.name, paste(c(confounders.names, interventions ), collapse=" + "), sep=" ~ "))

ca.stayhome <- prepDF(ca.chome, 
          policy, 
          states="CA",
          interventions,
          fips_codes)

# weekends filter
#ca.stayhome <- ca.stayhome%>% 
#      mutate(weekday= weekdays(as.Date(time_value)))%>% 
#      mutate(weekend= as.factor(ifelse(weekday %in% c("Saturday", "Sunday"), 1, 0)),
#             holiday = as.factor(ifelse(as.Date(time_value) %in% as.Date(ca.holidays), 1, 0)),
#             wildfire= as.factor(ifelse(as.Date(time_value) %in% as.Date(ca.wildfire_seasons), 1, 0)))



#interventions <- c("EmergDec","GathRestrict","SchoolClose","BarRestrict","PublicMask","OtherBusinessClose","BusinessMask", "SchoolMask" ,"Quarantine")

#formula1 <- as.formula(paste(mobility.name, paste(c(confounders.names, interventions ), collapse=" + "), sep=" ~ "))


# Filter the fips codes
filtered_fips <- fipscodes %>%filter(state %in% "CA")
ls <- list()
count <- 1
for(code in filtered_fips$fips){
  p <- ca.stayhome %>% filter(geo_value.x==code) %>%
    lm(formula1,data=.)
    ls[[count]] <- p 
  count <- count + 1 
  #print(fips_to_name(code))
  #print(summary(p))
}


```

## Completely stay at home: TX

```{r tx-stayathome}

interventions <- c("EmergDec",
                   "GathRestrict",
                   "SchoolClose",
                   "BarRestrict",
                   "PublicMask",
                   "OtherBusinessClose",
                   "Quarantine",
                   "StayAtHome")

mobility.name <- "completely_home_prop"

confounders.names <- c("confirmed_7dav_cumulative_prop",
                  "confirmed_7dav_incidence_prop",
                  "deaths_7dav_cumulative_prop",
                  "deaths_7dav_incidence_prop")

txformula <- as.formula(paste(mobility.name, paste(c(confounders.names, interventions), collapse=" + "), sep=" ~ "))

tx.stayhome <- prepDF(tx.chome, 
          policy, 
          states="TX",
          interventions,
          fips_codes)

# weekends filter
#tx.stayhome <- tx.stayhome%>% 
#      mutate(weekday= weekdays(as.Date(time_value)))%>% 
#      mutate(weekend= as.factor(ifelse(weekday %in% c("Saturday", "Sunday"), 1, 0)),
#             holiday = as.factor(ifelse(as.Date(time_value) %in% as.Date(tx.holidays), 1, 0)),
#             wildfire = as.factor(ifelse(as.Date(time_value) %in% as.Date(tx.wildfire_seasons), 1, 0)))


# interventions <- c("EmergDec",
#                    "GathRestrict",
#                    "SchoolClose",
#                    "BarRestrict",
#                    "PublicMask",
#                    "OtherBusinessClose",
#                    "Quarantine",
#                    "StayAtHome",
#                    "weekend",
#                    "holiday")
# 
# txformula <- as.formula(paste(mobility.name, paste(c(confounders.names, interventions), collapse=" + "), sep=" ~ "))


txls <- list()
count <- 1
# Filter the fips codes
tx_fips <- fipscodes %>%filter(state %in% "TX")
for(code in tx_fips$fips){
  p <- tx.stayhome %>% filter(geo_value.x==code) %>%
    lm(txformula,data=.)
  
  txls[[count]] <- p
  count <- count + 1
  #print(fips_to_name(code))
  #print(summary(p))
}

```


## Restaurant visit: CA

```{r ca-restaurant-visit}
interventions <- c("EmergDec",
                   "GathRestrict",
                   "SchoolClose",
                   "BarRestrict",
                   "PublicMask",
                   "OtherBusinessClose",
                   "BusinessMask",
                   "SchoolMask",
                   "Quarantine")

confounders.names <- c("confirmed_7dav_cumulative_prop",
                  "confirmed_7dav_incidence_prop",
                  "deaths_7dav_cumulative_prop",
                  "deaths_7dav_incidence_prop")

mobility.name <- "restaurants_visit_prop"


formula2 <- as.formula(paste(mobility.name, paste(c(confounders.names, interventions ), collapse=" + "), sep=" ~ "))

# CA restaurant visit
ca.restvisit <- prepDF(ca.rest, 
          policy, 
          states="CA",
          interventions,
          fips_codes)


# # weekends filter
# ca.restvisit <- ca.restvisit%>% 
#       mutate(weekday= weekdays(as.Date(time_value)))%>% 
#       mutate(weekend= as.factor(ifelse(weekday %in% c("Saturday", "Sunday"), 1, 0)),
#              holiday = as.factor(ifelse(as.Date(time_value) %in% as.Date(ca.holidays), 1, 0)))
# 
# 
# interventions <- c("EmergDec",
#                    "GathRestrict",
#                    "SchoolClose",
#                    "BarRestrict",
#                    "PublicMask",
#                    "OtherBusinessClose",
#                    "BusinessMask",
#                    "SchoolMask",
#                    "Quarantine",
#                    "weekend",
#                    "holiday")

formula2 <- as.formula(paste(mobility.name, paste(c(confounders.names, interventions ), collapse=" + "), sep=" ~ "))


# Filter the fips codes
filtered_fips <- fipscodes %>%filter(state %in% "CA")
ca.restls <- list()
ca.counties_filtered <- NULL
count <- 1
for(code in filtered_fips$fips){
  if(code %in% unique(ca.restvisit$geo_value.x)){
    p <- ca.restvisit %>% filter(geo_value.x==code) %>%
    lm(formula2,data=.)
    ca.restls[[count]] <- p
    countyname <- fips_to_name(code)
    ca.counties_filtered[count] <- countyname
    count <- count + 1 
    #print(countyname)
    #print(summary(p))
  }
}

# Show all results for county level mobility
#counter <- 1
#temp <- fips_to_name(unique(ca.restvisit$geo_value.x))
#for (i in ca.restls){
#  print(temp[counter])
#  print(summary(i))
#  counter <- counter + 1
#}


```

## Restaurant Visit: TX

```{r tx-restaurant-visit}
interventions <- c("EmergDec",
                   "GathRestrict",
                   "SchoolClose",
                   "BarRestrict",
                   "PublicMask",
                   "OtherBusinessClose",
                   "Quarantine")

confounders.names <- c("confirmed_7dav_cumulative_prop",
                  "confirmed_7dav_incidence_prop",
                  "deaths_7dav_cumulative_prop",
                  "deaths_7dav_incidence_prop")

mobility.name <- "restaurants_visit_prop"


formula2 <- as.formula(paste(mobility.name, paste(c(confounders.names, interventions ), collapse=" + "), sep=" ~ "))

# TX restaurant visit
tx.restvisit <- prepDF(tx.rest, 
          policy, 
          states="TX",
          interventions,
          fips_codes)


# weekends filter
# tx.restvisit <- tx.restvisit%>% 
#       mutate(weekday= weekdays(as.Date(time_value)))%>% 
#       mutate(weekend= as.factor(ifelse(weekday %in% c("Saturday", "Sunday"), 1, 0)),
#              holiday = as.factor(ifelse(as.Date(time_value) %in% as.Date(tx.holidays), 1, 0)))
# 
# interventions <- c("EmergDec",
#                    "GathRestrict",
#                    "SchoolClose",
#                    "BarRestrict",
#                    "PublicMask",
#                    "OtherBusinessClose",
#                    "Quarantine",
#                    "weekend",
#                    "holiday")
# formula2 <- as.formula(paste(mobility.name, paste(c(confounders.names, interventions ), collapse=" + "), sep=" ~ "))

# Filter the fips codes
filtered_fips <- fipscodes %>%filter(state %in% "TX")
tx.restls <- list()
count <- 1
tx.counties.filtered <- NULL
exclude_ls <- c("48391") # exclude this county due to lack of data
for(code in filtered_fips$fips){
  if(code %in% unique(tx.restvisit$geo_value.x) & !(code %in% exclude_ls)){
    p <- tx.restvisit %>% filter(geo_value.x==code) %>%
      lm(formula2,data=.)
    tx.restls[[count]] <- p
    countyname <- fips_to_name(code)
    tx.counties.filtered[count] <- countyname
    count <- count + 1
    # print all the information
    #print(countyname)
    #print(summary(p))
  }
}
```



# Rank of the effectiveness of the intervention

We can see that the effects of public mask and other business closures on reducing mobility are statistically significant at 0.05 level after we have counted for the case count signals. It also depends on the county regarding which intervention is more effective. For example, we see that school mask has a large effect on Stanislaus county, but may not have any effects on Alpine county. In addition, the interventions of gathering restriction and school closure in California are advisory, which appear to be not significant, whereas other mandatory policies have a significant effect on mobility.


## The effect of California's intervention on stay home signal

### CA: Urban counties

```{r ca-intervention-on-stay-home, fig.height= 12, fig.width = 10}
select_col <- c("EmergDec1","GathRestrict1","SchoolClose1","BarRestrict1","PublicMask1","OtherBusinessClose1","BusinessMask1", "SchoolMask1" ,"Quarantine1")
plist <- list()
count <- 1

for(i in 1:length(ls)){
  # Same the summary
  lm.fit <- summary(ls[[i]])
  name <- c("EmergDec","GathRestrict","SchoolClose","BarRestrict","PublicMask","OtherBusinessClose","BusinessMask", "SchoolMask" ,"Quarantine")
  df <- data.frame(fit= lm.fit$coefficients[select_col,1], interventions=name)
revised_df <- cbind(df,confint(ls[[i]])[select_col,])
  colnames(revised_df)[3] <- "LCI"
  colnames(revised_df)[4] <- "UCI"
  
  p <- plot_bar(revised_df,
                "fit",
                "interventions",
                "LCI",
                "UCI",
                "fit",
                ca.counties[i]) 
  plist[[count]] <- p
  count <- count + 1
}

  
# Plot all counties mobility signal
#n <- length(plist[c(2,19,41,53)])
#nCol <- floor(sqrt(n))
#do.call("grid.arrange", c(plist[c(2,19,41,53)], ncol=nCol))

idx <- which(ca.counties %in% c("Orange County", "Santa Clara County", "Alameda County", "Yolo County", "Marin County", "San Bernardino County"))

# Urban counties
do.call("grid.arrange", c(plist[idx], ncol=2))

```

### CA: Rural counties

```{r ca-rural-counties-on-stayhome, fig.height= 12, fig.width = 10}
# select counties
idx <- which(ca.counties %in% c("Humboldt County", "Siskiyou County", "Del Norte County", "Mono County", "Plumas County", "Glenn County"))
# Rural counties
do.call("grid.arrange", c(plist[idx], ncol=2))

```

## The effect of Texas's intervention on stay home signal

### TX: Urban counties

```{r tx-intervention-on-stay-home, fig.height= 12, fig.width = 10}
select_col <- c("EmergDec1","GathRestrict1","SchoolClose1","BarRestrict1","PublicMask1","Quarantine1","StayAtHome1")

plist <- list()
count <- 1
for(i in 1:length(txls)){
  # Same the summary
  lm.fit <- summary(txls[[i]])
  name <- c("EmergDec","GathRestrict","SchoolClose","BarRestrict","PublicMask","Quarantine","StayAtHome")
  df <- data.frame(fit= lm.fit$coefficients[select_col,1], interventions=name)
revised_df <- cbind(df,confint(txls[[i]])[select_col,])
  colnames(revised_df)[3] <- "LCI"
  colnames(revised_df)[4] <- "UCI"
  
  p <- plot_bar(revised_df,
                "fit",
                "interventions",
                "LCI",
                "UCI",
                "fit",
                tx.counties[i]) 
  plist[[count]] <- p
  count <- count + 1
}

# Plot all counties mobility signal
#nCol <- 1
#for(i in 1:length(plist)){
#  do.call("grid.arrange", c(plist[i], ncol=nCol))
#}


# Urban counties
idx <- which(fips_to_name(unique(tx.stayhome$geo_value.x)) %in% c("Chambers County","Hardin County"  ,"Harris County", "Fort Bend County","Liberty County","Austin County"))

do.call("grid.arrange", c(plist[idx], ncol=2))

```

### TX: Rural counties

```{r tx-urban-stay-home, fig.height= 12, fig.width = 10}

# Rural counties

idx <- which(fips_to_name(unique(tx.stayhome$geo_value.x)) %in%  c("Anderson County", "Andrews County", "Angelina County", "Somervell County", "Pecos County", "Ward County"))

do.call("grid.arrange", c(plist[idx], ncol=2))

```


## The effect of California's intervention on restaurant visit 

### CA: Urban counties

```{r ca-restaurant-rank, fig.height= 12, fig.width = 10}
select_col <- c("EmergDec1","GathRestrict1","SchoolClose1","BarRestrict1","PublicMask1","OtherBusinessClose1","BusinessMask1", "SchoolMask1" ,"Quarantine1")

plist <- list()
count <- 1
for(i in 1:length(ca.restls)){
  if(ca.counties_filtered[[i]]=="Sierra County"){
    next
  }
  # Same the summary
  lm.fit <- summary(ca.restls[[i]])
  name <- c("EmergDec","GathRestrict","SchoolClose","BarRestrict","PublicMask"        ,"OtherBusinessClose","BusinessMask", "SchoolMask" ,"Quarantine")
  df <- data.frame(fit= lm.fit$coefficients[select_col,1], interventions=name)
revised_df <- cbind(df,confint(ca.restls[[i]])[select_col,])
  colnames(revised_df)[3] <- "LCI"
  colnames(revised_df)[4] <- "UCI"
  
  p <- plot_bar(revised_df,
                "fit",
                "interventions",
                "LCI",
                "UCI",
                "fit",
                ca.counties_filtered[i]) 
  plist[[count]] <- p
  count <- count + 1
}

# Plot all counties mobility signal
#nCol <- 1
#for(i in 1:length(plist)){
#  do.call("grid.arrange", c(plist[i], ncol=nCol))
#}

idx <- which(fips_to_name(unique(ca.restvisit$geo_value.x)) %in% c("Orange County", "Santa Clara County", "Alameda County", "Yolo County", "Marin County", "San Bernardino County"))

# Urban counties
do.call("grid.arrange", c(plist[idx], ncol=2))

```

### CA: Rural counties

```{r ca-rural-counties-restaurant, fig.height= 12, fig.width = 10}
# select counties
idx <-which(fips_to_name(unique(ca.restvisit$geo_value.x)) %in% c("Humboldt County", "Siskiyou County", "Del Norte County", "Mono County", "Plumas County", "Glenn County"))
# Rural counties
do.call("grid.arrange", c(plist[idx], ncol=2))

```


## The effect of Texas's intervention on restaurant visit

### TX: Urban counties

```{r tx-intervention-on-restaurant, fig.height= 12, fig.width = 10}
select_col <- c("EmergDec1","GathRestrict1","SchoolClose1","BarRestrict1","PublicMask1","Quarantine1")

plist <- list()
count <- 1
for(i in 1:length(tx.restls)){
  # Same the summary
  lm.fit <- summary(tx.restls[[i]])
  name <- c("EmergDec","GathRestrict","SchoolClose","BarRestrict","PublicMask","Quarantine")
  df <- data.frame(fit= lm.fit$coefficients[select_col,1], interventions=name)
revised_df <- cbind(df,confint(tx.restls[[i]])[select_col,])
  colnames(revised_df)[3] <- "LCI"
  colnames(revised_df)[4] <- "UCI"
  
  p <- plot_bar(revised_df,
                "fit",
                "interventions",
                "LCI",
                "UCI",
                "fit",
                tx.counties.filtered[i]) 
  plist[[count]] <- p
  count <- count + 1
}

# Plot all counties mobility signal
#nCol <- 1
#for(i in 1:length(plist)){
#  do.call("grid.arrange", c(plist[i], ncol=nCol))
#}

# Urban counties
idx <- which(tx.counties.filtered %in% c("Chambers County","Hardin County"  ,"Harris County", "Fort Bend County","Liberty County","Austin County"))
do.call("grid.arrange", c(plist[idx], ncol=2))

```

### TX: Rural counties

```{r tx-rural-counties-restaurant, fig.height= 12, fig.width = 10}
# Rural counties

idx <- which(tx.counties.filtered %in% c("Anderson County", "Andrews County", "Angelina County", "Somervell County", "Pecos County", "Ward County"))
do.call("grid.arrange", c(plist[idx], ncol=2))
```



# Conclusion

We assume that every county will follow the state-wide policy accordingly. Having accounted for weekends and holidays, bar restriction and school clousure seem to have a greater effect on increasing staying at home signal among all other interventions in California. Other interventions, in contrary, reduce stay at home signal in California. The effects of the interventions vary across counties. On the other hand, emergency declaration increases the stay home signal in some counties of Texas. Also, in general, public mask reduces stay at home signal in Texas. 

Also, we see that wearing masks may help increase restaurant visit. For example, business mask seems to stimulate the increase of the restaurant visit. Moreover, public masks increases the number of restaurant visit in most of the counties in California. 

We rank the effectiveness of the intervention on mobility in terms of the regression coefficient in a linear regression setting. We leave characterization for the ranks of the effect of intervention on county-level as a future work. Other regression methods such as non-parametric regression should also be used for further study. 

