---
title: "Ranking the effect of intervention at countylevel"
author: "Kenneth Lee"
date: "06/01/2021"
output:
  html_document:
    code_folding: hide
    keep_md: yes
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Abstract

This notebook has two objectives. Firstly, it aims to measure the effect of interventions on mobility with other potential confounding signals such as confirmed case counts. The motivation is to observe whether the effect of intervention is greater than the effect of case signals on mobility. Secondly, we will rank the effects of various government interventions.

Due to the limitation of computation, we focus on analyzing the state of California and its counties. We will regress mobility signals on the covariates of interests and measure the effect of the covariates based on its regression coefficients.

```{r load_data, message=F, warning=F}
library(dplyr)
library(lubridate)
library(covidcast)
library(RcppRoll)
library(tidycensus)
library(reshape2)
library(zoo)
library(readxl)

source("code/loader.r")
source("code/parser.r")

# Load files 
confirmed_7dav_cumulative_prop <- read_excel("data/case_signals/ca-county-confirmed_7dav_cumulative_prop.xlsx")
 
confirmed_7dav_incidence_prop <- read_excel("data/case_signals/ca-county-confirmed_7dav_incidence_prop.xlsx")
 
deaths_7dav_cumulative_prop <- read_excel("data/case_signals/ca-county-deaths_7dav_cumulative_prop.xlsx")
# 
deaths_7dav_incidence_prop <- read_excel("data/case_signals/ca-county-deaths_7dav_incidence_prop.xlsx")

STARTDATE <- "2020-02-20"
ENDDATE <- "2021-01-03"

# Get fips codes
ca.fips_codes <- fips_codes %>% filter(state=="CA") %>%
  mutate(fips = paste(state_code, county_code , sep=""))


temp.ca.ftime <- read.csv("data/20200220-ca-ftime.csv",
                                       stringsAsFactors = F)

#temp.ca.ftime <- covidcast_signal(data_source = "safegraph", 
#                            signal ="full_time_work_prop",
#                            start_day = STARTDATE, 
#                            end_day = ENDDATE,
#                            geo_type = "county", 
#                            geo_values = ca.fips_codes$fips)


# preprocess data types
 confirmed_7dav_cumulative_prop$time_value <- as.Date(confirmed_7dav_cumulative_prop$time_value)
 
 confirmed_7dav_incidence_prop$time_value <- as.Date(confirmed_7dav_incidence_prop$time_value)
 
 deaths_7dav_cumulative_prop$time_value <- as.Date(deaths_7dav_cumulative_prop$time_value)
 
 deaths_7dav_incidence_prop$time_value <- as.Date(deaths_7dav_incidence_prop$time_value)
 
 temp.ca.ftime$time_value <- as.Date(temp.ca.ftime$time_value)
 
 confirmed_7dav_cumulative_prop$geo_value <- paste("0", confirmed_7dav_cumulative_prop$geo_value, sep="" )
 
 confirmed_7dav_incidence_prop$geo_value <- paste("0", confirmed_7dav_incidence_prop$geo_value, sep="" )
 
 deaths_7dav_cumulative_prop$geo_value <- paste("0", deaths_7dav_cumulative_prop$geo_value, sep="" )
 
 deaths_7dav_incidence_prop$geo_value <- paste("0", deaths_7dav_incidence_prop$geo_value, sep="" )
 
temp.ca.ftime$geo_value <- paste("0",temp.ca.ftime$geo_value, sep="")


# Policy data
policy <- load_policy()
ca.policy <- policy %>% filter(StatePostal == "ca", StateWide == 1)

# filter all counties
# California mobility signal
ca.ftime <- read.csv("data/ca.ftime.csv")
ca.counties <- as.character(unique(ca.ftime$county))


ca.policy_signal<- getSumOfPolicy(ca.policy , STARTDATE, ENDDATE)

# Turn the polical signal to be factors
factored.ca.policy.signal <- cbind(ca.policy_signal[1], 
                                lapply(ca.policy_signal[3:14],
                                       as.factor),
                                ca.policy_signal[15:16])


```

```{r helper-function, message=F, warning=F}
################ getForwardDays () ###################
# compute the number of days within a given day range that
# gives the highest spearman correlation between df1 and df2

# argument
# df1: a dataframe 
# df2: a dataframe that the data needs to be forwarded
# dt_vec: a vector contains a range of integers

getForwardDays <- function(df1, df2, dt_vec){
  # Empty list
  df_list <- vector("list", length(dt_vec))
  
  for (i in 1:length(dt_vec)) {
    df_list[[i]] <- covidcast_cor(df1, df2, dt_y = dt_vec[i],
                                  by = "geo_value", 
                                  method="spearman")

    df_list[[i]]$dt <- dt_vec[i]
  }
  
  # Stack into one big data frame
  df <- do.call(rbind, df_list)
  

  return(df)
}

```

# Lag analysis

We would like to forward mobility in time that is most correlated with the respective case count signals and use those case count signals in regression.

```{r ca-lag-analysis, message=F, warning=F}
# We want to select the day in which the case count is most correlated with the mobility in terms of spearman correlation in time

dt_vec = 0:100

# Confirmed cases
cum.confirmed.case.cor.df <- getForwardDays(temp.ca.ftime,confirmed_7dav_cumulative_prop, dt_vec)


daily.new.case.cor.df <- getForwardDays(temp.ca.ftime,confirmed_7dav_incidence_prop, dt_vec)

# Death case
cum.death.cor.df <- getForwardDays(temp.ca.ftime,deaths_7dav_cumulative_prop, dt_vec)

daily.death.case.cor.df <- getForwardDays(temp.ca.ftime,deaths_7dav_incidence_prop, dt_vec)

```


```{r shiftday, message=F, warning=F}
# Combine case and mobility
cum.confirmed.case <- confirmed_7dav_cumulative_prop[c("time_value","geo_value","value")]
new.confirmed.case <- confirmed_7dav_incidence_prop[c("time_value","geo_value","value")]
cum.death.case <- deaths_7dav_cumulative_prop[c("time_value","geo_value","value")]
new.death.case <- deaths_7dav_incidence_prop[c("time_value","geo_value","value")]
mobility <- temp.ca.ftime

# change name
colnames(mobility)[7] <- "full_time_work_prop"
colnames(cum.confirmed.case)[3] <- "confirmed_7dav_cumulative_prop"
colnames(new.confirmed.case)[3] <- "confirmed_7dav_incidence_prop"
colnames(cum.death.case)[3] <- "deaths_7dav_cumulative_prop"
colnames(new.death.case)[3] <- "deaths_7dav_incidence_prop"

# A list of confounders
confounders <- list(cum.confirmed.case, 
                    new.confirmed.case, 
                    cum.death.case,
                    new.death.case)

# Left join again with all other potential confounders
for (confounder in confounders){
  mobility <- left_join(mobility, confounder, by=c("time_value", "geo_value"))
}

# combine with policy signal
mobility <- left_join(mobility, factored.ca.policy.signal, by=c("time_value"))

```

## Model 1:

$$Y_t = \beta_{0}+ \sum_{i}\beta_{i}S_{i_{t}} + \sum_{k}\beta_{k}P_{k_{t-r}}$$, where $r$ varies depending on the county and the correlation between case signal and response

```{r model1, warning=F}
cof_ls <- list()
count <- 1

for(county in ca.fips_codes$fips){
  
  selected.df <- mobility %>% filter(geo_value==county)
  county.mobility <- mobility %>% filter(geo_value==county)
  
  # Get the days for most correlated
  
  # filter the data
  filtered.confirmed.case.df <- cum.confirmed.case.cor.df %>%
    filter(geo_value == county)
  
  filtered.daily.new.case.df <- daily.new.case.cor.df %>%
    filter(geo_value == county)
  
  filtered.cum.death.df <- cum.death.cor.df %>%
    filter(geo_value == county)
  
  filtered.daily.death.case.df <- daily.death.case.cor.df %>%
    filter(geo_value == county)
  
  # if there is no missing value, we go ahead to compute the switch time
  if(!any(is.na(filtered.confirmed.case.df$value))){
    
    days <- min(filtered.confirmed.case.df[abs(filtered.confirmed.case.df$value) == max(abs(filtered.confirmed.case.df$value)),]$dt)
    
    # switch the column 
    state<-county
    
    selected.df <- shiftDays(county.mobility, selected.df, days, c("confirmed_7dav_cumulative_prop"))
    
  }
  
  if(!any(is.na(filtered.daily.new.case.df$value))){
    
    days <- min(filtered.daily.new.case.df[abs(filtered.daily.new.case.df$value) == max(abs(filtered.daily.new.case.df$value)),]$dt)
      
    # switch the column 
    state<-county
    
    selected.df <- shiftDays(county.mobility, selected.df, days, c("confirmed_7dav_incidence_prop"))
    
  }
  
  if(!any(is.na(filtered.cum.death.df$value))){
    
    days <- min(filtered.cum.death.df[abs(filtered.cum.death.df$value) == max(abs(filtered.cum.death.df$value)),]$dt)
      
    # switch the column 
    state<-county
    
    selected.df <- shiftDays(county.mobility, selected.df, days, c("deaths_7dav_cumulative_prop"))
    
  }
  
  if(!any(is.na(filtered.daily.death.case.df$value))){
    
    days <- min(filtered.daily.death.case.df[abs(filtered.daily.death.case.df$value) == max(abs(filtered.daily.death.case.df$value)),]$dt)
      
    # switch the column 
    state<-county
    
    selected.df <- shiftDays(county.mobility, selected.df, days, c("deaths_7dav_incidence_prop"))
    
  }


  # fit the regression
   p <- selected.df  %>% 
     mutate(sumEmergDec = roll_sum(EmergDec, 2, align = "right", fill = NA),
            sumGathRestrict=roll_sum(GathRestrict, 2, align = "right", fill = NA),
            sumSchoolClose=roll_sum(SchoolClose, 2, align = "right", fill = NA),
            sumBarRestrict=roll_sum(BarRestrict, 2, align = "right", fill = NA),
            sumNEBusinessClose=roll_sum(NEBusinessClose, 2, align = "right", fill = NA),
            sumRestaurantRestrict=roll_sum(RestaurantRestrict, 2, align = "right", fill = NA),
            sumStayAtHome=roll_sum(StayAtHome, 2, align = "right", fill = NA),
            sumPublicMask=roll_sum(PublicMask, 2, align = "right", fill = NA),
            sumOtherBusinessClose=roll_sum(OtherBusinessClose, 2, align = "right", fill = NA),
            sumSchoolMask=roll_sum(SchoolMask, 2, align = "right", fill = NA),
            sumQuarantine=roll_sum(Quarantine, 2, align = "right", fill = NA),
            sumBusinessMask=roll_sum(BusinessMask, 2, align = "right", fill = NA)
            ) %>%
     lm(full_time_work_prop ~ confirmed_7dav_cumulative_prop+confirmed_7dav_incidence_prop+deaths_7dav_cumulative_prop+sumEmergDec + sumGathRestrict+sumSchoolClose+sumBarRestrict+sumNEBusinessClose+sumRestaurantRestrict+sumStayAtHome+sumPublicMask+sumOtherBusinessClose+sumBusinessMask+sumSchoolMask+sumQuarantine, data=.)
   
   # record the result
   cof_ls[[count]] <- p
   count <- count + 1
}

# Print all summaries
for(fit in cof_ls){
  print(summary(fit))
}
```

## Model 2:

$$Y_t = \beta_{0}+ \sum_{i}\beta_{i}S_{i_{t}} + \sum_{k}\beta_{k}P_{k_{t}}$$

```{r model2, warning=F}
mod2_cof_ls <- list()
count <- 1
for(county in ca.fips_codes$fips){
  p <- mobility%>%filter(geo_value==county)%>%
    lm(full_time_work_prop ~ confirmed_7dav_cumulative_prop+confirmed_7dav_incidence_prop+deaths_7dav_cumulative_prop+deaths_7dav_incidence_prop+EmergDec + GathRestrict+BarRestrict+NEBusinessClose+RestaurantRestrict+StayAtHome+PublicMask+OtherBusinessClose+BusinessMask+SchoolMask+Quarantine, data=.)
  mod2_cof_ls[[count]] <- p
  count <- count + 1
}

# Show all results for county level mobility
counter <- 1
for (i in mod2_cof_ls){
  print(ca.fips_codes$county[counter])
  print(summary(i))
  counter <- counter + 1
}

```

* We can also examine the effect when the weekends are droppped.

```{r model2-without-weekend}

```
