---
title: "Evaluating recommend and mandatory policy for Allegheny county"
author: "Kenneth Lee"
date: "24/01/2021"
output:
  html_document:
    code_folding: hide
    keep_md: yes
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Introduction

This report is to compare the impacts of recommended policy and mandatory policies on restaurant visit and staying at home in Alleghency county. We will use the following model:

$$Y_t = \beta_{0} + \sum_{i}\beta_{i}C_{i,t-1} + \sum_{k}\beta_{k}I_{k,t} \text{    for  } i = \{1,\dots,6\}, k = \{1,\dots,19\}$$

```{r import-packages, warning=FALSE, message=FALSE}
library(covidcast)
library(dplyr)
library(lubridate)
library(mgcv)
library(ggplot2)

source("code/loader.r")
source("code/parser.r")
```

```{r load-data, warning=FALSE, message=FALSE}
state <- "pa"
GEO_VALUE <- "42003" # Allegheny_County fip 
GEO_TYPE <- "county"
STARTDATE <- "2020-02-20" # Start from where we have case count signals
ENDDATE <- "2021-01-17" # fix a recent end date

# Read in county-level intervention data 
# from stanford crowdsourced data
allcountypolicy <- read.csv("data/stanford-county-level-intervention-data.csv")
# filter down to county of intersts
countypolicy <- allcountypolicy %>% filter(county == "Allegheny County")

# Read in state-wide policy data from UW
policy <- load_policy()
statepolicy <- policy %>% 
  filter(StatePostal==state, StateWide==1)


# Fetch case signals
dailyconfirmed <- covidcast_signal(data_source = "indicator-combination",
                                    signal ="confirmed_7dav_incidence_prop",
                                    start_day = STARTDATE, 
                                    end_day = ENDDATE,
                                    geo_type = GEO_TYPE, 
                                    geo_values = GEO_VALUE)

dailydeath <- covidcast_signal(data_source = "indicator-combination",
                                    signal ="deaths_7dav_incidence_prop",
                                    start_day = STARTDATE, 
                                    end_day = ENDDATE,
                                    geo_type = GEO_TYPE, 
                                    geo_values = GEO_VALUE)

cumconfirmed <- covidcast_signal(data_source = "indicator-combination",
                                    signal ="confirmed_7dav_cumulative_prop",
                                    start_day = STARTDATE, 
                                    end_day = ENDDATE,
                                    geo_type = GEO_TYPE, 
                                    geo_values = GEO_VALUE)

cumdeath <- covidcast_signal(data_source = "indicator-combination",
                                    signal ="deaths_7dav_cumulative_prop",
                                    start_day = STARTDATE, 
                                    end_day = ENDDATE,
                                    geo_type = GEO_TYPE, 
                                    geo_values = GEO_VALUE)

adjcli  <- covidcast_signal(data_source = "doctor-visits",
                                    signal ="smoothed_adj_cli",
                                    start_day = STARTDATE, 
                                    end_day = ENDDATE,
                                    geo_type = GEO_TYPE, 
                                    geo_values = GEO_VALUE)

smoothed_adj_covid19_from_claims <- covidcast_signal(data_source = "hospital-admissions",
                                    signal ="smoothed_adj_covid19_from_claims",
                                    start_day = STARTDATE, 
                                    end_day = ENDDATE,
                                    geo_type = GEO_TYPE, 
                                    geo_values = GEO_VALUE)


# Fetch mobility signals of interests 

STARTDATE <- as.Date(STARTDATE) + 1  # Forward the mobility one day in time
ENDDATE <- as.Date(ENDDATE) + 1

restvisit <-  covidcast_signal(data_source = "safegraph",
                                    signal ="restaurants_visit_prop",
                                    start_day = STARTDATE, 
                                    end_day = ENDDATE,
                                    geo_type = GEO_TYPE, 
                                    geo_values = GEO_VALUE)

barvisit <-  covidcast_signal(data_source = "safegraph",
                                    signal ="bars_visit_prop",
                                    start_day = STARTDATE, 
                                    end_day = ENDDATE,
                                    geo_type = GEO_TYPE, 
                                    geo_values = GEO_VALUE)



```

```{r helper-functions, warning=FALSE, message=FALSE}
plot_bar <- function(df,
                     x,
                     y,
                     lower.bound,
                     upper.bound,
                     gp,
                     name){
  p<- ggplot(df, aes_string(x=x, y=y, group=gp, color=gp)) + 
  #geom_line() +
  geom_point()+
  geom_errorbar(aes_string(xmin=lower.bound, xmax=upper.bound), width=.2,
                 position=position_dodge(0.05))+
    ggtitle(name)
      
  return(p)
}

# process each policy as follows
# 1. look for the policy enact time, call it start date
# 2. then look for the policy ended time
# 3. if ended time is NA, look for expiry date, if not, set it as end date
# 4. if expiry date is NA, then it means the policy hasn't ended, if not, set it as enddate
# 5. Form a sequence of start date and end dates, assign 1 to any dates fall into this sequence.
concatCaseNPolicy <- function(case.df){
  # Loop through each row in the column called policy
  for(row in (1:nrow(statepolicy))){
    # for each policy, we do
    policy.name <- statepolicy[row,]$StatePolicy
    modified.policyName <- statepolicy[row,]$policy

    # Set the start date of the policy
    start <- statepolicy[row,]$DateEnacted
    
    # Set the end date of the policy
    if(is.na(statepolicy[row,]$DateEnded)){
      # if DateEnded is NA, then look for DateExpiry
      if(is.na(statepolicy[row,]$DateExpiry)){
        # if DateExpiry is NA, then the policy is in effect until the next policy with the same name happens
        
        # check to see if the same policy exist
        if(row == nrow(statepolicy)){
          # If this is the last policy in the data,
          # take the last day of the case signal
          end <- lubridate::today()
        }else{
          # if it is not the last policy, then look for the next one 
          
          # check if we still have the same policy
          if(policy.name %in% statepolicy[row+1:nrow(statepolicy),]$StatePolicy){
            # if so, we look for the first one
            for(k in (row+1:nrow(statepolicy))){
              if(statepolicy[k,]$StatePolicy==policy.name){
                if(statepolicy[k,]$DateEnacted == start){
                   end <- lubridate::today()
                }else{
                  end <- statepolicy[k,]$DateEnacted-1
                }
                break
              }
            }
          }else{
            # if not, then we will assign 
            end <- lubridate::today()
          }
        }
      }else{
        end <- statepolicy[row,]$DateExpiry - 1
      }
    }else{
      end <- statepolicy[row,]$DateEnded - 1
    }
    
    policy.period <- seq(start, end, 1)
    
    assigned <- ifelse(case.df$time_value %in% policy.period, 1, 0)
    
    # check if there are any duplicate names
    if(modified.policyName %in% colnames(case.df)){
      # if so, we assign 1's to that created column

      case.df[, modified.policyName] <- case.df[, modified.policyName] + assigned
    }else{
      # if not, create a new column and assign one to that column within the time
      case.df[, modified.policyName] <- assigned
    }
  }
  return(case.df)
}


```


```{r preprocess-case-signals}
# Filter the dates that we don't have mobility signals
missingdates <- c(seq(as.Date("2020-12-13"),as.Date("2020-12-19"), 1),
                  as.Date("2021-01-17"))

# Filter the case signals
dailyconfirmed.f <- dailyconfirmed %>% 
  filter(!(time_value %in% missingdates)) %>%
  mutate(dailyconfirmed_prop = value)

dailydeath.f <- dailydeath %>% 
  filter(!(time_value %in% missingdates)) %>%
  mutate(dailydeath_prop = value)

cumconfirmed.f <- cumconfirmed %>% 
  filter(!(time_value %in% missingdates)) %>%
  mutate(cumconfirmed_prop = value) 

cumdeath.f <- cumdeath %>% 
  filter(!(time_value %in% missingdates)) %>%
  mutate(cumdeath_prop = value) 

adjcli.f <- adjcli %>% 
  filter(!(time_value %in% missingdates)) %>%
  mutate(smoothed_adj_cli = value ) 

hospital.f<- smoothed_adj_covid19_from_claims %>%
  filter(!(time_value %in% missingdates)) %>%
  mutate(smoothed_adj_covid19_from_claims  = value ) 

# Concatnate case counts and 
restvisit$dailyconfirmed_prop <- dailyconfirmed.f$dailyconfirmed_prop
restvisit$dailydeath_prop <- dailydeath.f$dailydeath_prop
restvisit$cumconfirmed_prop <- cumconfirmed.f$cumconfirmed_prop 
restvisit$cumdeath_prop <- cumdeath.f$cumdeath_prop
restvisit$smoothed_adj_cli <-adjcli.f$smoothed_adj_cli
restvisit$smoothed_adj_covid19_from_claims <- hospital.f$smoothed_adj_covid19_from_claims

colnames(restvisit)[7] <- "restaurants_visit_prop"

barvisit$dailyconfirmed_prop <- dailyconfirmed.f$dailyconfirmed_prop
barvisit$dailydeath_prop <- dailydeath.f$dailydeath_prop
barvisit$cumconfirmed_prop <- cumconfirmed.f$cumconfirmed_prop 
barvisit$cumdeath_prop <- cumdeath.f$cumdeath_prop
barvisit$smoothed_adj_cli <-adjcli.f$smoothed_adj_cli 
barvisit$smoothed_adj_covid19_from_claims <- hospital.f$smoothed_adj_covid19_from_claims

colnames(barvisit)[7] <- "bars_visit_prop"

```

```{r preprocess-policy}
# Combine some information for gathering restriction

# get all indicies that are gathering restriction
gathering.idx <- which(statepolicy$StatePolicy == "GathRestrict")

statepolicy$policy <- paste(statepolicy$StatePolicy,statepolicy$Mandate, sep="")

limit.idx <- paste(statepolicy$InGathLim,
      statepolicy$OutGathLim,
      statepolicy$InGathLimReligious,
      statepolicy$OutGathLimReligious,
      sep="-")

# Get all the indicies wrt gathering restriction
statepolicy[gathering.idx,]$policy <- paste(limit.idx[gathering.idx],statepolicy[gathering.idx,]$policy , sep="")


lmdf <- concatCaseNPolicy(restvisit)

# convert to factor
lmdf[,c(17:ncol(lmdf))] <- sapply(lmdf[,c(17:ncol(lmdf))], function(x) as.factor(x)) 

```

## Restaurant visit prop

Having accounted for the potential confounding variables in terms of case count, we can see that manadatory stay home order has significantly reduced restaurant visit among all the interventions. Secondly, gathering restriction tends to be more effective as the gathering limit decreases. 

```{r analysis-parametric-restvisit}
# Covaraites: all case signals t-1, all interventions, ease indicator, expand indicator
# reponse: restaurant visit/ bar visits
df <- lmdf[,c(7, c(11:ncol(lmdf)))]
names <- colnames(df)[-1]

lm.fit <- lm(restaurants_visit_prop ~ ., df)
sum.lm.fit <- summary(lm.fit)

plt.df <- data.frame(fit= sum.lm.fit$coefficients[c(-1:-7),1], interventions=names(sum.lm.fit$coefficients[c(-1:-7),1]))

revised_df <- cbind(plt.df, confint(lm.fit)[which(names(lm.fit$coefficients) %in% names(sum.lm.fit$coefficients[c(-1:-7),1])),])

colnames(revised_df)[3] <- "LCI"
colnames(revised_df)[4] <- "UCI"
  
sum.lm.fit

plot_bar(revised_df,
              "fit",
              "interventions",
              "LCI",
              "UCI",
              "fit",
               "Allegheny County") 
```

## Bar visit prop

We can see a similar pattern as we change the response variable to bar visit per 100,000 population. However, only mandatory stay at home order's regression coefficient does not contain 0, meaning that it is statistically siginificant at 0.05 level that stay-at-home order reduces bar visit.

```{r analysis-parametric-barvisit}
barlmdf <- concatCaseNPolicy(barvisit)
# convert to factor
barlmdf[,c(17:ncol(barlmdf))] <- sapply(lmdf[,c(17:ncol(lmdf))], function(x) as.factor(x)) 

# Covaraites: all case signals t-1, all interventions, ease indicator, expand indicator
# reponse: restaurant visit/ bar visits
df <- barlmdf[,c(7, c(11:ncol(barlmdf)))]
names <- colnames(df)[-1]

lm.fit <- lm(bars_visit_prop ~ ., df)
sum.lm.fit <- summary(lm.fit)

plt.df <- data.frame(fit= sum.lm.fit$coefficients[c(-1:-7),1], interventions=names(sum.lm.fit$coefficients[c(-1:-7),1]))

revised_df <- cbind(plt.df, confint(lm.fit)[which(names(lm.fit$coefficients) %in% names(sum.lm.fit$coefficients[c(-1:-7),1])),])

colnames(revised_df)[3] <- "LCI"
colnames(revised_df)[4] <- "UCI"
  
sum.lm.fit

plot_bar(revised_df,
              "fit",
              "interventions",
              "LCI",
              "UCI",
              "fit",
               "Allegheny County") 
```

