---
title: "Evaluating recommend and mandatory policy for Allegheny county"
author: "Kenneth Lee"
date: "24/01/2021"
output:
  html_document:
    code_folding: hide
    keep_md: yes
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Introduction

This report is to compare the impacts of recommended policy and mandatory policies on restaurant visit and staying at home in Alleghency county. We will use the following model:

$$Y_t = \beta_{0} + \sum_{i}\beta_{i}C_{i,t-1} + \sum_{k}\beta_{k}I_{k,t} \text{    for  } i = \{1,\dots,6\}, k = \{1,\dots,19\}$$



```{r import-packages, warning=FALSE, message=FALSE}
library(covidcast)
library(dplyr)
library(lubridate)
library(mgcv)
library(ggplot2)
library(readxl)
library(gridExtra)

source("code/loader.r")
source("code/parser.r")
```

```{r load-data, warning=FALSE, message=FALSE}
state <- "pa"
GEO_VALUE <- "42003" # Allegheny_County fip 
GEO_TYPE <- "county"
STARTDATE <- "2020-02-20" # Start from where we have case count signals
ENDDATE <- "2021-01-17" # fix a recent end date

# Read in county-level intervention data 
# from stanford crowdsourced data
allcountypolicy <- read.csv("data/stanford-county-level-intervention-data.csv")
# filter down to county of intersts
countypolicy <- allcountypolicy %>% filter(county == "Allegheny County")

# Read in state-wide policy data from UW
policy <- load_policy()
statepolicy <- policy %>% 
  filter(StatePostal==state, StateWide==1)


# Fetch case signals
dailyconfirmed <- covidcast_signal(data_source = "indicator-combination",
                                    signal ="confirmed_7dav_incidence_prop",
                                    start_day = STARTDATE, 
                                    end_day = ENDDATE,
                                    geo_type = GEO_TYPE, 
                                    geo_values = GEO_VALUE)

dailydeath <- covidcast_signal(data_source = "indicator-combination",
                                    signal ="deaths_7dav_incidence_prop",
                                    start_day = STARTDATE, 
                                    end_day = ENDDATE,
                                    geo_type = GEO_TYPE, 
                                    geo_values = GEO_VALUE)

# Fetch mobility signals of interests 

STARTDATE <- as.Date(STARTDATE) + 1  # Forward the mobility one day in time
ENDDATE <- as.Date(ENDDATE) + 1

restvisit <-  covidcast_signal(data_source = "safegraph",
                                    signal ="restaurants_visit_prop",
                                    start_day = STARTDATE, 
                                    end_day = ENDDATE,
                                    geo_type = GEO_TYPE, 
                                    geo_values = GEO_VALUE)

barvisit <-  covidcast_signal(data_source = "safegraph",
                                    signal ="bars_visit_prop",
                                    start_day = STARTDATE, 
                                    end_day = ENDDATE,
                                    geo_type = GEO_TYPE, 
                                    geo_values = GEO_VALUE)


chome <- covidcast_signal(data_source = "safegraph", 
                             signal ="completely_home_prop",
                             start_day = STARTDATE, 
                             end_day = ENDDATE,
                             geo_type = GEO_TYPE, 
                             geo_values = GEO_VALUE)
```

```{r helper-functions, warning=FALSE, message=FALSE}
plot_bar <- function(df,
                     x,
                     y,
                     lower.bound,
                     upper.bound,
                     gp,
                     name){
  p<- ggplot(df, aes_string(x=x, y=y, group=gp, color=gp)) + 
  #geom_line() +
  geom_point()+
  geom_errorbar(aes_string(xmin=lower.bound, xmax=upper.bound), width=.2,
                 position=position_dodge(0.05))+
    ggtitle(name)
      
  return(p)
}

# process each policy as follows
# 1. look for the policy enact time, call it start date
# 2. then look for the policy ended time
# 3. if ended time is NA, look for expiry date, if not, set it as end date
# 4. if expiry date is NA, then it means the policy hasn't ended, if not, set it as enddate
# 5. Form a sequence of start date and end dates, assign 1 to any dates fall into this sequence.
concatCaseNPolicy <- function(case.df){
  # Loop through each row in the column called policy
  for(row in (1:nrow(statepolicy))){
    # for each policy, we do
    policy.name <- statepolicy[row,]$StatePolicy
    modified.policyName <- statepolicy[row,]$policy

    # Set the start date of the policy
    start <- statepolicy[row,]$DateEnacted
    
    # Set the end date of the policy
    if(is.na(statepolicy[row,]$DateEnded)){
      # if DateEnded is NA, then look for DateExpiry
      if(is.na(statepolicy[row,]$DateExpiry)){
        # if DateExpiry is NA, then the policy is in effect until the next policy with the same name happens
        
        # check to see if the same policy exist
        if(row == nrow(statepolicy)){
          # If this is the last policy in the data,
          # take the last day of the case signal
          end <- lubridate::today()
        }else{
          # if it is not the last policy, then look for the next one 
          
          # check if we still have the same policy
          if(policy.name %in% statepolicy[row+1:nrow(statepolicy),]$StatePolicy){
            # if so, we look for the first one
            for(k in (row+1:nrow(statepolicy))){
              if(statepolicy[k,]$StatePolicy==policy.name){
                if(statepolicy[k,]$DateEnacted == start){
                   end <- lubridate::today()
                }else{
                  end <- statepolicy[k,]$DateEnacted-1
                }
                break
              }
            }
          }else{
            # if not, then we will assign 
            end <- lubridate::today()
          }
        }
      }else{
        end <- statepolicy[row,]$DateExpiry - 1
      }
    }else{
      end <- statepolicy[row,]$DateEnded - 1
    }
    
    policy.period <- seq(start, end, 1)
    
    assigned <- ifelse(case.df$time_value %in% policy.period, 1, 0)
    
    # check if there are any duplicate names
    if(modified.policyName %in% colnames(case.df)){
      # if so, we assign 1's to that created column

      case.df[, modified.policyName] <- case.df[, modified.policyName] + assigned
    }else{
      # if not, create a new column and assign one to that column within the time
      case.df[, modified.policyName] <- assigned
    }
  }
  return(case.df)
}


```

```{r preprocess-temperature}
patemp <- read_excel("data/temperature/patemp.xlsx")
patemp$DATE <- as.Date(patemp$DATE)
patemp.df <- patemp %>% select(DATE, PRCP, TAVG)
colnames(patemp.df) <- c("time_value", "precipitation", "Avg.temperature")

```



```{r preprocess-case-signals}
# Add data to chome
chome$dailyconfirmed_prop <- dailyconfirmed$value
chome$dailydeath_prop <- dailydeath$value
colnames(chome)[7] <- "completely_home_prop"

# Filter the dates that we don't have mobility signals
missingdates <- c(seq(as.Date("2020-12-13"),as.Date("2020-12-19"), 1),
                  as.Date("2021-01-17"))

# Filter the case signals
dailyconfirmed.f <- dailyconfirmed %>% 
  filter(!(time_value %in% missingdates)) %>%
  mutate(dailyconfirmed_prop = value)

dailydeath.f <- dailydeath %>% 
  filter(!(time_value %in% missingdates)) %>%
  mutate(dailydeath_prop = value)

# Concatnate case counts and 
restvisit$dailyconfirmed_prop <- dailyconfirmed.f$dailyconfirmed_prop
restvisit$dailydeath_prop <- dailydeath.f$dailydeath_prop

colnames(restvisit)[7] <- "restaurants_visit_prop"

barvisit$dailyconfirmed_prop <- dailyconfirmed.f$dailyconfirmed_prop
barvisit$dailydeath_prop <- dailydeath.f$dailydeath_prop

colnames(barvisit)[7] <- "bars_visit_prop"

```

```{r preprocess-policy}
# Combine some information for gathering restriction

# get all indicies that are gathering restriction
gathering.idx <- which(statepolicy$StatePolicy == "GathRestrict")

statepolicy$policy <- paste(statepolicy$StatePolicy,statepolicy$Mandate, sep="")

limit.idx <- paste(statepolicy$InGathLim,
      statepolicy$OutGathLim,
      statepolicy$InGathLimReligious,
      statepolicy$OutGathLimReligious,
      sep="-")

# Get all the indicies wrt gathering restriction
statepolicy[gathering.idx,]$policy <- paste(limit.idx[gathering.idx],statepolicy[gathering.idx,]$policy , sep="")



   
```

## Restaurant visit prop

Having accounted for the potential confounding variables in terms of case count, we can see that manadatory stay home order has significantly reduced restaurant visit among all the interventions. Secondly, gathering restriction tends to be more effective as the gathering limit decreases. 

```{r plot-restvisit-with-weekends}
# Plot the signal
p <- ggplot(restvisit, aes(x=time_value, y=restaurants_visit_prop)) +
  geom_line() + 
  labs(title = "Number of daily restaurant visit by time in Allegaheny county (weekends included)", y= "# of resturant visit per 100,000 population")
p

```

### With weekends

```{r analysis-parametric-restvisit}
restvisit.temp <- left_join(restvisit, patemp.df, by ="time_value")

lmdf <- concatCaseNPolicy(restvisit.temp)

# Add holidays
county.holidays <- as.Date(c("2020-01-01", 
                 "2020-01-20",
                 "2020-02-17",
                 "2020-04-28",
                 "2020-05-25",
                 "2020-07-03",
                 "2020-09-07",
                 "2020-10-12",
                 "2020-11-11",
                 "2020-11-26",
                 "2020-12-25",
                 "2021-01-01",
                 "2021-01-18"))

# Drop all weekends
lmdf <- lmdf %>% 
      mutate(holiday = as.factor(ifelse(time_value %in% county.holidays, 1, 0)))

colnames(lmdf) <- make.names(colnames(lmdf))
# convert to factor
lmdf[,c(15:ncol(lmdf))] <- sapply(lmdf[,c(15:ncol(lmdf))], function(x) as.factor(x)) 


# Covaraites: all case signals t-1, all interventions, ease indicator, expand indicator
# reponse: restaurant visit/ bar visits
df <- lmdf[,c(7, c(11:ncol(lmdf)))]
names <- colnames(df)[-1]

lm.fit <- lm(restaurants_visit_prop ~ ., df)
sum.lm.fit <- summary(lm.fit)

plt.df <- data.frame(fit= sum.lm.fit$coefficients[c(-1:-5, -25),1], interventions=names(sum.lm.fit$coefficients[c(-1:-5, -25),1]))

revised_df <- cbind(plt.df, confint(lm.fit)[which(names(lm.fit$coefficients) %in% names(sum.lm.fit$coefficients[c(-1:-5, -25),1])),])

colnames(revised_df)[3] <- "LCI"
colnames(revised_df)[4] <- "UCI"
  
sum.lm.fit

plot_bar(revised_df,
              "fit",
              "interventions",
              "LCI",
              "UCI",
              "fit",
               "Allegheny County; Restaurant visit; Weekends included") 
```


### Without weekends

```{r plot-restvisit-without-weekends}
lmdf <- concatCaseNPolicy(restvisit.temp)

# Drop all weekends
lmdf <- lmdf %>% 
      mutate(weekday= weekdays(as.Date(time_value)),
             holiday = as.factor(ifelse(time_value %in% county.holidays, 1, 0)))%>% 
      filter(!(weekday %in% c("Saturday", "Sunday"))) 
   
# Drop the weekday variable
lmdf <- lmdf[, !(colnames(lmdf) %in% c("weekday"))]



# Plot the signal
p <- ggplot(lmdf, aes(x=time_value, y=restaurants_visit_prop)) +
  geom_line() + 
  labs(title = "Number of daily restaurant visit by time in Allegaheny county (weekends dropped)", y= "# of resturant visit per 100,000 population")
p

# Plot the interventions
plist <- list()
count <- 1

colnames(lmdf) <- make.names(colnames(lmdf))
for(i in 15:(ncol(lmdf)-1)){
  name <- colnames(lmdf)[i]
  plist[[count]] <- ggplot(lmdf, aes_string(x="time_value", y=name)) +
    geom_line() + 
  labs(title = name, y= "In effect or not")
  
  count <- count + 1
}

# Plot all the plots
do.call("grid.arrange", c(plist[c(1:5)], ncol=1))
do.call("grid.arrange", c(plist[c(6:10)], ncol=1))
do.call("grid.arrange", c(plist[c(11:15)], ncol=1))
do.call("grid.arrange", c(plist[c(16:20)], ncol=1))
do.call("grid.arrange", c(plist[c(21:22)], ncol=1))


# convert to factor
lmdf[,c(15:ncol(lmdf))] <- sapply(lmdf[,c(15:ncol(lmdf))], function(x) as.factor(x)) 


# Covaraites: all case signals t-1, all interventions, ease indicator, expand indicator
# reponse: restaurant visit/ bar visits
df <- lmdf[,c(7, c(11:ncol(lmdf)))]
names <- colnames(df)[-1]

lm.fit <- lm(restaurants_visit_prop ~ ., df)
sum.lm.fit <- summary(lm.fit)

plt.df <- data.frame(fit= sum.lm.fit$coefficients[c(-1:-5, -25),1], interventions=names(sum.lm.fit$coefficients[c(-1:-5, -25),1]))

revised_df <- cbind(plt.df, confint(lm.fit)[which(names(lm.fit$coefficients) %in% names(sum.lm.fit$coefficients[c(-1:-5, -25),1])),])

colnames(revised_df)[3] <- "LCI"
colnames(revised_df)[4] <- "UCI"
  
sum.lm.fit

plot_bar(revised_df,
              "fit",
              "interventions",
              "LCI",
              "UCI",
              "fit",
               "Allegheny County; Restaurant visit; Weekends dropped") 
```



## Bar visit prop

We can see a similar pattern as we change the response variable to bar visit per 100,000 population. However, only mandatory stay at home order's regression coefficient does not contain 0, meaning that it is statistically siginificant at 0.05 level that stay-at-home order reduces bar visit.

### With weekends

```{r plot-bar-withweekends}
# Plot the signal
p <- ggplot(barvisit, aes(x=time_value, y=bars_visit_prop)) +
  geom_line() + 
  labs(title = "Number of daily bars visit by time in Allegaheny county (weekends included)", y= "# of bars visit per 100,000 population")
p
```


```{r analysis-parametric-barvisit-with-weekends}
barvisit.df <- left_join(barvisit, patemp.df, by ="time_value")
barlmdf <- concatCaseNPolicy(barvisit.df)

# Add holidays
barlmdf <- barlmdf %>% 
      mutate(holiday = as.factor(ifelse(time_value %in% county.holidays, 1, 0)))

colnames(barlmdf) <- make.names(colnames(barlmdf))

# convert to factor
barlmdf[,c(15:ncol(barlmdf))] <- sapply(barlmdf[,c(15:ncol(barlmdf))], function(x) as.factor(x)) 


# Covaraites: all case signals t-1, all interventions, ease indicator, expand indicator
# reponse: restaurant visit/ bar visits
df <- barlmdf[,c(7, c(11:ncol(barlmdf)))]
names <- colnames(df)[-1]

lm.fit <- lm(bars_visit_prop ~ ., df)
sum.lm.fit <- summary(lm.fit)

plt.df <- data.frame(fit= sum.lm.fit$coefficients[c(-1:-5, -25),1], interventions=names(sum.lm.fit$coefficients[c(-1:-5, -25),1]))

revised_df <- cbind(plt.df, confint(lm.fit)[which(names(lm.fit$coefficients) %in% names(sum.lm.fit$coefficients[c(-1:-5, -25),1])),])

colnames(revised_df)[3] <- "LCI"
colnames(revised_df)[4] <- "UCI"
  
sum.lm.fit

plot_bar(revised_df,
              "fit",
              "interventions",
              "LCI",
              "UCI",
              "fit",
               "Allegheny County; Bars visit; Weekends included") 
```

### Without weekends

```{r analysis-parametric-barvisit-without-weekends}
barvisit.df <- left_join(barvisit, patemp.df, by ="time_value")
barlmdf <- concatCaseNPolicy(barvisit.df)

# Drop all weekends
barlmdf <- barlmdf %>% 
      mutate(weekday= weekdays(as.Date(time_value)),
             holiday = as.factor(ifelse(time_value %in% county.holidays, 1, 0)))%>% 
      filter(!(weekday %in% c("Saturday", "Sunday"))) 

# Drop the weekday variable
barlmdf <- barlmdf[, !(colnames(barlmdf)%in% c("weekday"))]

# Plot one without weekends
p <- ggplot(barlmdf, aes(x=time_value, y=bars_visit_prop)) +
  geom_line() + 
  labs(title = "Number of daily bars visit by time in Allegaheny county (weekends dropped)", y= "# of bars visit per 100,000 population")
p

colnames(barlmdf) <- make.names(colnames(barlmdf))
# convert to factor
barlmdf[,c(15:ncol(barlmdf))] <- sapply(barlmdf[,c(15:ncol(barlmdf))], function(x) as.factor(x)) 


# Covaraites: all case signals t-1, all interventions, ease indicator, expand indicator
# reponse: restaurant visit/ bar visits
df <- barlmdf[,c(7, c(11:ncol(barlmdf)))]
names <- colnames(df)[-1]

lm.fit <- lm(bars_visit_prop ~ ., df)
sum.lm.fit <- summary(lm.fit)

plt.df <- data.frame(fit= sum.lm.fit$coefficients[c(-1:-5, -25),1], interventions=names(sum.lm.fit$coefficients[c(-1:-5, -25),1]))

revised_df <- cbind(plt.df, confint(lm.fit)[which(names(lm.fit$coefficients) %in% names(sum.lm.fit$coefficients[c(-1:-5, -25),1])),])

colnames(revised_df)[3] <- "LCI"
colnames(revised_df)[4] <- "UCI"
  
sum.lm.fit

plot_bar(revised_df,
              "fit",
              "interventions",
              "LCI",
              "UCI",
              "fit",
               "Allegheny County; Bars visit; Weekends dropped") 

```

## Completely staying home

### With weekends

```{r chome-analysis-with-weekend}
chome.df <- left_join(chome, patemp.df, by ="time_value")
chomedf <- concatCaseNPolicy(chome.df)

# Add holidays
chomedf <- chomedf %>% 
      mutate(holiday = as.factor(ifelse(time_value %in% county.holidays, 1, 0)))

colnames(chomedf) <- make.names(colnames(chomedf))

# convert to factor
chomedf[,c(15:ncol(chomedf))] <- sapply(chomedf[,c(15:ncol(chomedf))], function(x) as.factor(x)) 


# Covaraites: all case signals t-1, all interventions, ease indicator, expand indicator
# reponse: restaurant visit/ bar visits
df <- chomedf[,c(7, c(11:ncol(chomedf)))]
names <- colnames(df)[-1]

lm.fit <- lm(completely_home_prop ~ ., df)
sum.lm.fit <- summary(lm.fit)

plt.df <- data.frame(fit= sum.lm.fit$coefficients[c(-1:-5, -25),1], interventions=names(sum.lm.fit$coefficients[c(-1:-5, -25),1]))

revised_df <- cbind(plt.df, confint(lm.fit)[which(names(lm.fit$coefficients) %in% names(sum.lm.fit$coefficients[c(-1:-5, -25),1])),])

colnames(revised_df)[3] <- "LCI"
colnames(revised_df)[4] <- "UCI"
  
sum.lm.fit

plot_bar(revised_df,
              "fit",
              "interventions",
              "LCI",
              "UCI",
              "fit",
               "Allegheny County; Completely Staying Home; With weekends") 

```

### Without weekends

```{r}
chome.df <- left_join(chome, patemp.df, by ="time_value")
chomedf <- concatCaseNPolicy(chome.df)

# Add holidays
chomedf <- chomedf %>% 
      mutate(weekday= weekdays(as.Date(time_value)),
             holiday = as.factor(ifelse(time_value %in% county.holidays, 1, 0)))%>% 
      filter(!(weekday %in% c("Saturday", "Sunday"))) 


# Drop the weekday variable
chomedf<- chomedf[, !(colnames(chomedf)%in% c("weekday"))]

colnames(chomedf) <- make.names(colnames(chomedf))
# convert to factor
chomedf[,c(15:ncol(chomedf))] <- sapply(chomedf[,c(15:ncol(chomedf))], function(x) as.factor(x)) 


# Covaraites: all case signals t-1, all interventions, ease indicator, expand indicator
# reponse: restaurant visit/ bar visits
df <- chomedf[,c(7, c(11:ncol(chomedf)))]
names <- colnames(df)[-1]

lm.fit <- lm(completely_home_prop ~ ., df)
sum.lm.fit <- summary(lm.fit)

plt.df <- data.frame(fit= sum.lm.fit$coefficients[c(-1:-5, -25),1], interventions=names(sum.lm.fit$coefficients[c(-1:-5, -25),1]))

revised_df <- cbind(plt.df, confint(lm.fit)[which(names(lm.fit$coefficients) %in% names(sum.lm.fit$coefficients[c(-1:-5, -25),1])),])

colnames(revised_df)[3] <- "LCI"
colnames(revised_df)[4] <- "UCI"
  
sum.lm.fit

plot_bar(revised_df,
              "fit",
              "interventions",
              "LCI",
              "UCI",
              "fit",
               "Allegheny County; Completely Staying Home; Weekends dropped") 
```