---
title: "Indicator Heterogeneity I: Temporal Heterogeneity"
author: "Addison"
output:
  html_document
---
```{r import_statements, echo = FALSE, message = FALSE}
library(tidyverse)
library(covidcast)
```

# Background

Notebook to play with slope coefficients before putting them into the report.

# Data setup

```{r data_ingestion_county, echo = TRUE, cache=TRUE}
# Fetch the following sources and signals from the API 
# TODO: Add Google Symptoms "eventually"
source_names = c("doctor-visits", "fb-survey", "fb-survey", "hospital-admissions")
signal_names = c("smoothed_adj_cli", "smoothed_cli", "smoothed_hh_cmnty_cli", 
            "smoothed_adj_covid19")
pretty_names = c("Doctor visits", "Facebook CLI", "Facebook CLI-in-community", 
          "Hospitalizations")
target_names = c("Cases", "Cases", "Cases", "Deaths")
geo_level = "county"

start_day = "2020-04-15"
end_day = NULL
cache_fname = 'cached_data/03_heterogeneity_core_indicators.RDS'

if (!file.exists(cache_fname)) {
  df_signals = vector("list", length(signal_names))
  for (i in 1:length(signal_names)) {
    df_signals[[i]] = suppressWarnings(
                        covidcast_signal(source_names[i], signal_names[i],
                                         start_day, end_day,
                                         geo_type=geo_level))
  }

  # Fetch USAFacts confirmed case incidence proportion (smoothed with 7-day 
  # trailing average)
  df_cases = suppressWarnings(
              covidcast_signal("usa-facts", "confirmed_7dav_incidence_prop",
                              start_day, end_day,
                              geo_type=geo_level))

  df_deaths = suppressWarnings(
              covidcast_signal("usa-facts", "deaths_7dav_incidence_prop",
                              start_day, end_day,
                              geo_type=geo_level))

  case_num = 500
  geo_values = suppressWarnings(covidcast_signal("usa-facts", "confirmed_cumulative_num",
                                max(df_cases$time_value), 
                                max(df_cases$time_value))) %>%
    filter(value >= case_num) %>% pull(geo_value)
  saveRDS(list(df_signals, df_cases, df_deaths), cache_fname)
} else {
  cached_data = readRDS(cache_fname)
  df_signals = cached_data[[1]]
  df_cases = cached_data[[2]]
  df_deaths = cached_data[[3]]
}
```

## Setup

```{r plot_slopes, echo = TRUE, cache=TRUE}
sensorize_time_ranges = list(
      c(-7, -1),
      c(-10, -1),
      c(-14, -1),
      c(-21, -1))
QUANTS = c(0.01, 0.99)

# TODO: Add more "core indicators"

for (ind_idx in 1:length(source_names)) {
  if (target_names[ind_idx] == 'Cases') {
    df_target = df_cases
  } else if (target_names[ind_idx] == 'Deaths') {
    df_target = df_deaths
  } else {
    stop(sprintf("No matching dataframe for target %s.", target_names[ind_idx]))
  }

  base_cor_fname = sprintf('results/03_base_cors_%s_%s.RDS',
                            source_names[ind_idx], signal_names[ind_idx])
  sensorize_fname = sprintf('results/03_sensorize_cors_%s_%s.RDS',
                            source_names[ind_idx], signal_names[ind_idx])
  sensorize_val_fname = sprintf('results/03_sensorize_vals_%s_%s.RDS',
                            source_names[ind_idx], signal_names[ind_idx])

  df_cor_base = readRDS(base_cor_fname)
  sensorize_cors = readRDS(sensorize_fname)
  sensorized_vals = readRDS(sensorize_val_fname)

  for (inner_idx in 1:length(sensorize_time_ranges)) {
    sv = sensorized_vals[[inner_idx]]
    print(summary(sv$slope))
    print(slope_limits <- quantile(sv$slope, QUANTS, na.rm=TRUE))
    plt = ggplot(
      sensorized_vals[[inner_idx]],
      aes(x=time_value,
          y=slope),
    ) + geom_point (
      alpha=0.1,
      size=0.5,
    ) + geom_hline (
      yintercept=0,
      colour='white',
    ) + stat_summary (
        aes(y=slope,
            group=1,
            colour='median'),
        fun=median,
        geom="line",
        group=1,
    ) + stat_summary (
        aes(y=slope,
            group=1,
            colour='+/- mad'),
        fun=function(x) { median(x) + mad(x) },
        geom="line",
        group=1,
    ) + stat_summary (
        aes(y=slope,
            group=1,
            colour='+/- mad'),
        fun=function(x) { median(x) - mad(x) },
        geom="line",
        group=1,
    ) + scale_colour_manual(
        values=c("median"="maroon",
                 "+/- mad"="darkgreen")
    ) + labs(
      colour=''
    ) + ggtitle(
      sprintf("Slope distribution for %s, fitted on t in %d:%d",
              pretty_names[ind_idx],
              sensorize_time_ranges[[inner_idx]][1],
              sensorize_time_ranges[[inner_idx]][2])
    ) + ylim (
      slope_limits[[1]], slope_limits[[2]]
    )
    print(plt)
  }
}

```

Why is the distribution of slope coefficients centered on zero?  (Shouldn't we
expect the center of the distribution to be positive?)  It _might_ be possible
that the distribution is not centered on zero once we condition on location
(but unlikely).  We can take a look anyways...

```{r condition_on_location, echo = TRUE}
sensorize_time_ranges = list(
      c(-7, -1),
      c(-10, -1),
      c(-14, -1),
      c(-21, -1))
QUANTS = c(0.01, 0.99)

# TODO: Add more "core indicators"

for (ind_idx in 1:length(source_names)) {
  if (target_names[ind_idx] == 'Cases') {
    df_target = df_cases
  } else if (target_names[ind_idx] == 'Deaths') {
    df_target = df_deaths
  } else {
    stop(sprintf("No matching dataframe for target %s.", target_names[ind_idx]))
  }

  base_cor_fname = sprintf('results/03_base_cors_%s_%s.RDS',
                            source_names[ind_idx], signal_names[ind_idx])
  sensorize_fname = sprintf('results/03_sensorize_cors_%s_%s.RDS',
                            source_names[ind_idx], signal_names[ind_idx])
  sensorize_val_fname = sprintf('results/03_sensorize_vals_%s_%s.RDS',
                            source_names[ind_idx], signal_names[ind_idx])

  df_cor_base = readRDS(base_cor_fname)
  sensorize_cors = readRDS(sensorize_fname)
  sensorized_vals = readRDS(sensorize_val_fname)

  for (inner_idx in 1:length(sensorize_time_ranges)) {
    sv = sensorized_vals[[inner_idx]]
    sv_medmad = sv %>% group_by(
      geo_value
    ) %>% summarize (
      med = median(slope),
      mad = mad(slope),
      med_upper = med+mad,
      med_lower = med-mad,
    )
    coverage = sv_medmad %>% mutate (
      contains_zero = (0 < med_upper) & (0 > med_lower),
    ) %>% pull (
      contains_zero
    ) %>% mean (
      na.rm=TRUE
    )
    print(sprintf('Coverage for %s: %f',
                  pretty_names[ind_idx],
                  coverage))
  }
}

```
