---
title: "Estimating Causal Effect of Government Intervention to Mobility via Regression Discontinuity"
author: "Kenneth Lee"
date: "23/09/2020"
output: 
  html_document:
    code_folding: hide
    keep_md: yes
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

```{r import packages, echo=FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library (readr)
library(tidyverse)
library(dplyr)
library(covidcast)
library(lubridate)
library(ggpubr)
library(reshape2)
library(tidyr)
library(viridis)
library(gridExtra)
library(data.table)
library(zoo)
library(cowplot)

source("code/painter.r")
source("code/loader.r")
source("code/parser.r")
```

```{r define global variables}
STARTDATE <- "2019-01-01"
ENDDATE <- lubridate::today()
GEO_TYPE = "state" # state-level
GEO_VALUE = "ca" # all states
EXCLUDED_AREAS = c("as","gu", "mp","vi") # excluded areas due to small sample size

# Pick one of the following to specify mobility signal from delphi API
# "full_time_work_prop"       "part_time_work_prop"      
# "completely_home_prop"     "median_home_dwell_time"  

MOBILITY <- "full_time_work_prop"


# Pick one of the following to specify policy
# Please refer to data source code book for details

# "EmergDec"            "SchoolClose"        
# "GathRestrict"        "BarRestrict"        
# "OtherBusinessClose"  "RestaurantRestrict" 
# "CaseIsolation"       "StayAtHome"         
# "PublicMask"          "Quarantine"         
# "NEBusinessClose"     "TravelRestrictIntra"
# "TravelRestrictEntry" "SchoolMask"         
# "TravelRestrictExit"  "BusinessMask"   

POLICY <- "EmergDec"
```

```{r import data, warning = FALSE, message=FALSE, echo=FALSE}
# Get the covidcast data accordingly
data <- load_covidcast_data(STARTDATE, 
                            ENDDATE, 
                            GEO_TYPE, 
                            GEO_VALUE, 
                            EXCLUDED_AREAS)

policy <- load_policy()
intervention <- getSumOfPolicy(policy, STARTDATE, ENDDATE)
```

# Full-time away home signal

This notebook only looks at California state with a particular intervention named emergency declaration

```{r preprocess, warning = FALSE, message=FALSE, fig.height= 15, fig.width = 15}

# Filter to only specified one state
all.policies <- intervention %>%
  filter(geo_value == GEO_VALUE)

# We fist get all the signals
mobility <- data[[MOBILITY]][c("time_value","value")] #mobility

case.count <- data$Avg.Confirmed.Case.Count[c("time_value",
                                              "value")]# case count

cumulative.case.count <-data$Cum.Avg.Case.Count[c("time_value",
                                                  "value")]# cumulative case count

cumulative.case.count.prop <-data$Cum.Avg.Case.Count.Prop[c("time_value",
                                                            "value")] # cumulative case count per 100,000

avg.death.count <- data$Avg.Death.Case.Count[c("time_value",
                                               "value")] #death count

cum.avg.death.count <- data$Cum.Avg.Death.Count[c("time_value",
                                                  "value")] #cumulative death count

doc.visit <- data$smoothed_cli[c("time_value",
                                 "value")] # doctor visit

adj.doc.visit <- data$smoothed_adj_cli[c("time_value",
                                         "value")] # adjusted doctor visit

# Instead of calling "value", we distingush the value by its original name
colnames(mobility)[2]<- MOBILITY
colnames(case.count)[2] <- "confirmed_7dav_incidence_num"
colnames(cumulative.case.count)[2] <- "confirmed_7dav_cumulative"
colnames(cumulative.case.count.prop)[2] <- "confirmed_7dav_cumulative_prop"
colnames(avg.death.count)[2]<-"deaths_7dav_incidence_num"
colnames(cum.avg.death.count)[2] <- "deaths_7dav_cumulative_num"
colnames(doc.visit)[2] <- "smoothed_cli"
colnames(adj.doc.visit)[2] <- "smoothed_adj_cli"

# Create a list of confounders for left join with mobility
confounders <- list(case.count, 
                    cumulative.case.count, 
                    cumulative.case.count.prop,
                    avg.death.count, 
                    cum.avg.death.count, 
                    doc.visit, 
                    adj.doc.visit)

# Left join mobility dataframe with policy dataframe in CA
intervention_mobility_case <- joinMobilityAndPolicy(mobility,
                                                    all.policies,
                                                    confounders)

# Get the first time stamp of the intervention in the specified state
intervention.first.day <- getFirstDayOfIntervention(intervention, GEO_VALUE, POLICY)

# Plot the time-series: mobility signal as y, time as x 
plotInterventionTime(intervention_mobility_case, 
                                mobility.name=MOBILITY,
                                intervention.first.day,
                                ylab= MOBILITY,
                                xlab="time-value")
```

## Lag analysis at a state-level

```{r state-lag-analysis, echo=FALSE, warning = FALSE, message=FALSE, fig.height= 15, fig.width = 15}

SHIFTDAY <- 100
corr.method <- "spearman"
by_method <- "geo_value"
title <- "Median Rank correlation between other signals and future mobility (slicing by state)"

covidcastlike.signals <- list(
  data$Avg.Confirmed.Case.Count, 
  data$Cum.Avg.Case.Count, 
  data$Cum.Avg.Case.Count.Prop, 
  data$Avg.Death.Case.Count, 
  data$Cum.Avg.Death.Count, 
  data$smoothed_cli, 
  data$smoothed_adj_cli)

names <- list("7-day avg. confirmed case", 
             "Cum 7day avg. confirmed case",
             "Cum 7day avg. confirmed case, per 100,000",
             "death case",
             "cumulative death case",
             "doctor visit",
             "doctor visit (day-of-week effects removed)")


# Compute pearson correlation between other covidcast-like signals and mobility
plot.all.Corr.Median.by.shift(covidcastlike.signals, 
                              data[[MOBILITY]],
                              SHIFTDAY,
                              names, 
                              corr.method, 
                              title, 
                              by_method)
```

## Use most correlated shift time to build linear models

$$y = \beta_{0} + \beta_{1}(\sum I(A_i = 1)) + \beta_{2}x_1 + \beta_{3}x_2$$


```{r linear regression}
# select best lag number based on intial exploration
case.shifted.days.spearman <- 37 # based spearman correlation
doc.visit.shifted.days.spearman <- 37 # based spearman correlation for doctor visit
cum.death.shifted.days.spearman <- 55
death.shifted.days.spearman <- 42

case.vec <- c("confirmed_7dav_incidence_num", "confirmed_7dav_cumulative", "confirmed_7dav_cumulative_prop")
doc.vec <- c("smoothed_cli", "smoothed_adj_cli")
cum.death.vec <- c("deaths_7dav_cumulative_num")
death.vec <- c("deaths_7dav_incidence_num")

# Make two copies for shifting the data
selected.df <- intervention_mobility_case
factored_data <- intervention_mobility_case
# Change the data by shifting the covariates
# Shift the case count column vector by the specified shift time
factored_data <- shiftDays(selected.df, factored_data, case.shifted.days.spearman, case.vec)

# Shift the  doctor column vector by the specified shift time
factored_data <-shiftDays(selected.df, factored_data, doc.visit.shifted.days.spearman, doc.vec)
  
# Shift the death count column vector by the specified shift time
factored_data <-shiftDays(selected.df, factored_data, cum.death.shifted.days.spearman, cum.death.vec)

# Shift the cum death count column vector by the specified shift time
factored_data <-shiftDays(selected.df, factored_data, death.shifted.days.spearman, death.vec)
  
variables <- c("intervention.duration", "smoothed_cli","smoothed_adj_cli")
formula <- as.formula(
  paste(MOBILITY, 
        paste(variables, collapse = " + "), 
        sep = " ~ "))

# We specifically look at emergency declaration
dummy.df <- factored_data%>% 
  mutate(intervention.duration = cumsum(eval(parse(text=POLICY))))

lm(formula=formula,dummy.df) %>% summary()
```

```{r more linear models}
# We specifically look at emergency declaration

variables <- c("intervention.duration",
               "smoothed_cli",
               "smoothed_adj_cli", 
               "confirmed_7dav_cumulative_prop",
               "deaths_7dav_cumulative_num")

formula <- as.formula(
  paste(MOBILITY, 
        paste(variables, collapse = " + "), 
        sep = " ~ "))

dummy.df <- factored_data%>% 
  mutate(intervention.duration = cumsum(eval(parse(text=POLICY)))) 

lm(formula, dummy.df) %>% 
  summary()
```

## Model with time only (weekends included)

```{r time-only regression-weekends-included, echo=FALSE, warning = FALSE, message=FALSE, fig.height= 15, fig.width = 15}
formula <- as.formula(paste(MOBILITY, "time_value", sep = " ~ "))

# Fit a linear regression model
lm.fit <- lm(formula, data =factored_data)

# Pad the fitted values with NA 
pred.y <- predict(lm.fit)

# Plot the graph
p <- factored_data %>% 
  mutate(Intervention = as.factor(eval(parse(text=POLICY)))) %>% 
  ggplot(aes_string(x = "time_value", y = MOBILITY, color = "Intervention")) +
  geom_point() + 
  geom_line(aes(x = time_value, 
                y = pred.y, 
                colour="fitted value"), 
            size = 1) +
   labs(title = paste(MOBILITY,
                      "regressing on time (weekends included)", 
                      sep=" "))
p
```

## Model with shifting the covariates (weekends included)

```{r plot the regression, echo=FALSE, warning = FALSE, message=FALSE, fig.height= 15, fig.width = 15}
# Create a new dataframe
factored_data.lm <- factored_data%>% 
  mutate(EmergDec.duration = cumsum(EmergDec),
         StayAtHome.duration = cumsum(StayAtHome),
         PublicMask.duration = cumsum(PublicMask),
         SchoolClose.duration = cumsum(SchoolClose),
         GathRestrict.duration = cumsum(GathRestrict),
         TravelRestrictExit.duration = cumsum(TravelRestrictExit),
         TravelRestrictEntry.duration = cumsum(TravelRestrictEntry),
         TravelRestrictIntra.duration = cumsum(TravelRestrictIntra),
         BarRestrict.duration = cumsum(BarRestrict),
         NEBusinessClose.duration = cumsum(NEBusinessClose),
         Quarantine.duration = cumsum(Quarantine),
         RestaurantRestrict.duration = cumsum(RestaurantRestrict),
         CaseIsolation.duration = cumsum(CaseIsolation))

# Form the formula
intervention_char <- paste(POLICY, "duration", sep=".")

variables <- c(intervention_char,
               "smoothed_cli",
               "smoothed_adj_cli")

formula <- as.formula(
  paste(MOBILITY, 
        paste(variables, collapse = " + "), 
        sep = " ~ "))


# Fit a linear regression model
lm.fit <- lm(formula, data =factored_data.lm)  

############# TO-DO : FIX THE NUMBER ##############

# Pad the fitted values with NA 
factored_data.lm$predlm <- c(rep(NA, 433),predict(lm.fit))

# Plot the graph
p <- factored_data.lm %>% 
  mutate(policy.duration = cumsum(eval(parse(text=POLICY))), Intervention = as.factor(eval(parse(text=POLICY)))) %>% 
  ggplot(aes_string(x = "time_value", y = MOBILITY, color = "Intervention")) +
  geom_point() + 
  geom_line(aes(x = time_value, y = predlm, colour="fitted value"), size = 1) +
   labs(title = "Covariates selected WITH most correlated number of shift")
p

```

## Model without shifting the covariates (weekends included)

```{r LR model without selecting most correlated lag covariates, warning = FALSE, message=FALSE, fig.height= 15, fig.width = 15}

variables <- c("intervention.duration",
               "smoothed_cli",
               "smoothed_adj_cli")

formula <- as.formula(
  paste(MOBILITY, 
        paste(variables, collapse = " + "), 
        sep = " ~ "))

intervention.lm <- intervention_mobility_case %>% 
  mutate(intervention.duration = cumsum(eval(parse(text=POLICY)))) 

lm.fit.no.lag <- lm(formula, data =intervention.lm) 

intervention.lm$predlm <- c(rep(NA, nrow(intervention.lm) - length(predict(lm.fit.no.lag))), predict(lm.fit.no.lag))

intervention.lm%>% 
  mutate(policy.duration = cumsum(eval(parse(text=POLICY))), 
         Intervention = as.factor(eval(parse(text=POLICY)))) %>% 
  ggplot(aes_string(x = "time_value", y = MOBILITY, color ="Intervention")) +
  geom_point() + 
  geom_line(aes(x = time_value, y = predlm, colour="fitted value"), size = 1)+
  labs(title = "Covariates selected WITHOUT most correlated number of shift")
```

## Weekend effects

We suspect that the mobility signal is lower than usual during the weekend.

```{r examine weekend effects from the data}
intervention_mobility_case$weekday <- weekdays(as.Date(intervention_mobility_case$time_value)) 

p <- ggplot(intervention_mobility_case, 
            aes_string(x="weekday", y=MOBILITY)) + 
  geom_boxplot()
p
```


## Re-plot the regression line after dropping weekends

## Regressing on time only

```{r time-only-regression-weekends-excluded, echo=FALSE, warning = FALSE, message=FALSE, fig.height= 15, fig.width = 15}
# Get lm summary to compare
factored_data.without.weekend <- factored_data.lm%>% 
  mutate(Intervention = as.factor(eval(parse(text=POLICY))), 
         weekday= weekdays(as.Date(time_value)))%>% 
  filter(!weekday %in% c("Saturday", "Sunday")) 

formula <- as.formula(paste(MOBILITY, "time_value", sep=" ~ "))

# Fit a linear regression model
lm.fit <- lm(formula , data =factored_data.without.weekend)

# Pad the fitted values with NA 
pred.y <- predict(lm.fit)

# Plot the graph
p <- factored_data.without.weekend %>% 
  mutate(Intervention = as.factor(eval(parse(text=POLICY)))) %>% 
  ggplot(aes_string(x = "time_value", 
             y = MOBILITY, 
             color = "Intervention")) +
  geom_point() + 
  geom_line(aes(x = time_value, 
                y = pred.y, 
                colour="fitted value"), 
            size = 1) +
   labs(title = paste(MOBILITY,"regressing on time (weekends excluded)" , sep= " "))
p
```

### Covariates shifted

```{r plot the regression line for the data WITH shift after dropping weekends, echo=FALSE, warning = FALSE, message=FALSE, fig.height= 15, fig.width = 15}

# Form the formula
intervention_char <- paste(POLICY, "duration", sep=".")

formula.head <- paste(MOBILITY, 
                      intervention_char,
                      sep=" ~ ")

form.string <- paste(formula.head,
      "smoothed_cli",
      "smoothed_adj_cli",
      sep="+")
      
formula <- as.formula(form.string)

# Check the lm.fit
lm(formula,data=factored_data.without.weekend) %>% 
  summary()

# Get all the fitted values
shifted.predlm <- lm(formula,data=factored_data.without.weekend) %>% predict()

# Pad the fitted values with NA 
factored_data.without.weekend$predlm <- c(rep(NA, nrow(factored_data.without.weekend) - length(shifted.predlm)), shifted.predlm)

factored_data.without.weekend %>%
  ggplot(aes_string(x = "time_value", 
                    y = MOBILITY, 
                    color = "Intervention")) +
  geom_point() + 
  geom_line(aes(y = predlm, 
                colour="fitted value"), 
            size = 1) +
   labs(title = "Covariates selected WITH most correlated number of shift (weekends dropped)")

```


### Covariates not shifted


```{r plot the regression line for the data WITHOUT shift after dropping weekends, echo=FALSE, warning = FALSE, message=FALSE, fig.height= 15, fig.width = 15}
# Filter the dataset
noshift.without.weekends <-intervention_mobility_case%>%
  mutate(intervention.duration = cumsum(eval(parse(text=POLICY))), 
         Intervention = as.factor(eval(parse(text=POLICY))), 
         weekday = weekdays(as.Date(time_value))) %>% 
  filter(!weekday %in% c("Saturday", "Sunday"))
  
# Form the formula
formula.head <- paste(MOBILITY, 
                      "intervention.duration",
                      sep=" ~ ")

form.string <- paste(formula.head,
      "smoothed_cli",
      "smoothed_adj_cli",
      sep="+")
      
formula <- as.formula(form.string)

# Print lm summary
lm(formula, noshift.without.weekends) %>% summary()
  
# Get the fitted value  
noshift.predlm <- lm(formula, noshift.without.weekends) %>% predict()

# Pad the fitted values with NA 
noshift.without.weekends$predlm <- c(rep(NA, nrow(noshift.without.weekends) - length(noshift.predlm)), noshift.predlm)

# Plot the graph
noshift.without.weekends %>%
  ggplot(aes_string(x = "time_value", 
                    y = MOBILITY, 
                    color = "Intervention")) +
  geom_point() + 
  geom_line(aes(y = predlm, 
                colour="fitted value"), 
            size = 1)+
  labs(title = "Covariates selected WITHOUT most correlated number of shift (weekends dropped)")
```

### What will happen if we add more covidcast signals as covariates?

#### Adding covidcast signals

```{r more covariates to without weekend models,echo=FALSE, warning = FALSE, message=FALSE, fig.height= 15, fig.width = 15}
# Add more confounders to the regression model
# Check the lm.fit
factored_data.without.weekend %$%
  lm(full_time_work_prop ~ EmergDec.duration + smoothed_cli+smoothed_adj_cli+confirmed_7dav_incidence_num+confirmed_7dav_cumulative+confirmed_7dav_cumulative_prop+deaths_7dav_incidence_num+deaths_7dav_cumulative_num) %>% 
  summary()

# Get all the fitted values
shifted.predlm <- factored_data.without.weekend%$%
  lm(full_time_work_prop ~ EmergDec.duration + smoothed_cli+smoothed_adj_cli+confirmed_7dav_incidence_num+confirmed_7dav_cumulative+confirmed_7dav_cumulative_prop+deaths_7dav_incidence_num+deaths_7dav_cumulative_num) %>%
  predict()

# Pad the fitted values with NA 
factored_data.without.weekend$predlm <- c(rep(NA, nrow(factored_data.without.weekend) - length(shifted.predlm)), shifted.predlm)

factored_data.without.weekend %>%
  ggplot(aes_string(x = "time_value", y = MOBILITY, color = "Intervention")) +
  geom_point() + 
  geom_line(aes(y = predlm, colour="fitted value"), size = 1) +
   labs(title = "All covariates selected WITH most correlated number of shift (weekends dropped)")

```

#### Adding other policy duration

```{r add more other policy duration,warning = FALSE, message=FALSE, fig.height= 15, fig.width = 15}
# Try to add other intervention covariates
formula.head <- paste(MOBILITY, 
                 "EmergDec.duration", sep=" ~ ")

formula <- as.formula(paste(formula.head,
      "StayAtHome.duration",
      "PublicMask.duration",
      "SchoolClose.duration",
      "GathRestrict.duration",
      "BarRestrict.duration",
      "NEBusinessClose.duration",
      "RestaurantRestrict.duration",
      "smoothed_cli",
      "smoothed_adj_cli",
      "confirmed_7dav_incidence_num",
      "confirmed_7dav_cumulative",
      "confirmed_7dav_cumulative_prop",
      "deaths_7dav_incidence_num",
      "deaths_7dav_cumulative_num",
      "SchoolClose.duration", sep="+"))


lm(formula, factored_data.without.weekend) %>% 
  summary()

# Predict the mobility
new.pred <- lm(formula, factored_data.without.weekend) %>% predict()

# Pad the fitted values with NA 
factored_data.without.weekend$predlm <- c(rep(NA, nrow(factored_data.without.weekend) - length(new.pred)), new.pred)

# Plot the graph
factored_data.without.weekend %>%
  ggplot(aes_string(x = "time_value", 
                    y = MOBILITY,
                    color = "Intervention")) +
  geom_point() + 
  geom_line(aes(y = predlm, 
                colour="fitted value"), size = 1) +
   labs(title = "All covariates selected WITH most correlated number of shift (weekends dropped)")

```


